---
title: "Diet.description"
output: html_document
---

#Load Libraries
```{r}
library(dplyr)
library(readr)
library(Rpath)
library(tidyverse)

GroupInfo_CalCur_USE <- read_csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/GroupInfo_CalCur-USE.csv")

Diet <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Diet_CalCur_USE.csv")  
Diet <- Diet %>% rename("Juv. hake" = "Juv..hake", "Yellowtail rock."= "Yellowtail.rock.", "Shelf rock."= "Shelf.rock.", "P. Ocean Perch"= "P..Ocean.Perch", "Longspine thorny."= "Longspine.thorny.", "Tufted Puffin"= "Tufted.Puffin", "Leach's S. Petrel"= "Leach.s.S..Petrel","Fin whale"= "Fin.whale", "Slope rock."= "Slope.rock.", "carniv-zoops"= "carniv.zoops", "small jellies"= "small.jellies", "large jellies"= "large.jellies", "pandalid shp"= "pandalid.shp", "benthic shp"= "benthic.shp", "tanner crab"= "tanner.crb", "Benthic fish"= "Benthic.fish", "surf perch"= "surf.perch", "Other cephal."= "Other.cephal.", "Market squid"= "Market.squid", "Pacific Mackerel"= "Pacific.Mackerel")
Diet
```


#Define datasets
```{r}
predators_of_mesopelagic <- Diet[,c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]
mesopelagic_pred <- Diet[15, c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]
prey_of_mesopelagic <- Diet[c(3,6:10,15,18), c(1,16)]
forage.pred <- Diet[19:27, c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]
euphausiids <- Diet[7, c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]

species <- as.data.frame(forage.pred$DIETS) %>% mutate(X=1:9) %>% rename("Species"= "forage.pred$DIETS")
#Check if this is right 
forage.pred$MAX <- apply(forage.pred[,-1],1,max)
forage <- as.data.frame(forage.pred$MAX) %>% mutate(X=1:9)
forage <- forage %>% rename("Max diet"= "forage.pred$MAX") %>% left_join(species, by= "X") %>% select(-X)
forage
mesopelagic.pred$MAX <- apply(mesopelagic.pred[,-1],1,max,na.rm=TRUE)
mesopelagic <- as.data.frame(mesopelagic.pred$MAX) %>% rename("Max diet"= "mesopelagic.pred$MAX") %>% mutate("Species"= "mesopelagic fish")
euphausiids$MAX <- apply(euphausiids[,-1],1,max,na.rm=TRUE)
euphausiids <- as.data.frame(euphausiids$MAX) %>% rename("Max diet"= "euphausiids$MAX") %>% mutate("Species"= "euphausiids")

combination <- bind_rows(forage, mesopelagic, euphausiids)

predators_of_mesopelagic[15,]
```

#Predators of mesopelagic fish: diet dependence on mesopelagics and forage fish
```{r}
#For each predator of mesopelagic fish find their prey items that make up at least 5% of their diet
predators_of_mesopelagic

 #Webplot highlighting predator and prey connections to mesopelagic fish  
jpeg(file="~/Documents/OTZ.Manuscript/Images/Sub.2/web.plot.jpeg",
width=300, height=525)
webplot(Cal_bal, eco.name= attr(Cal_bal, "California Current"), line.col="grey", labels=TRUE, highlight= "mesopelagics", highlight.col= c("blue", "red", "NA"), label.num=TRUE, label.cex=0.7, fleets=FALSE)

tiff("~/Documents/OTZ.Manuscript/Images/Sub.2/web.plot.jpeg", units="in", width=6, height=5, res=600, compression = 'lzw')
webplot(Cal_bal, eco.name= attr(Cal_bal, "California Current"), line.col="grey", labels=FALSE, highlight= "mesopelagics", highlight.col= c("blue", "red", "NA"), label.num=TRUE, label.cex=0.7, fleets=FALSE) + standard_theme
dev.off()
```

#Grouped graph: weighted aggregate biomass for predators of mesopelagic fish, other prey of predators of mesopelagic fish (not just prey of mesopelagic fish)
```{r}
data_1 <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/data.1.csv")
data_2 <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/data.2.csv")

#Define a common ggplot theme 
standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.text.align= 0, legend.title= element_text(size = 10), legend.text = element_text(size= 9), axis.text=element_text(size=9), axis.title=element_text(size=10))
 
#Graph 1: predators of mesopelagic fish, mesopelagic fish, other prey of predators of mesopelagic fish

#Pred_mesopelagic_prey_0.05 is made up of the predators of mesopelagic fish that have other prey consisting of at least 5% of their diet, predators_mesopelagic_0.05 are the predators of mesopelagic fish that have at least 5% of their diet consist of mesopelagics 
pred_mesopelagic_prey_0.05 <- predators_of_mesopelagic %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% select(DIETS) %>% unique() %>% rename(spp = DIETS) %>% mutate(classification= "Other prey of predators of mesopelagics") 

#Remove mesopelagics as predator of mesopelagics, not included as non-commercial or commercial predator in the study
predators_mesopelagic_0.05 <- predators_of_mesopelagic[15,] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% select(name) %>% mutate(classification= "Predators of mesopelagics") %>% rename(spp= name)
combo_g1 <- rbind(pred_mesopelagic_prey_0.05, predators_mesopelagic_0.05)
combo_g1 <-combo_g1[!duplicated(combo_g1$spp),] 
GroupInfo_edt <- GroupInfo_CalCur_USE %>% select(Group, Biomass) %>% rename(spp= Group) #select original biomass from group info 
data_2_g1 <- data_2 %>% left_join(GroupInfo_edt, by= "spp") %>% left_join(combo_g1, by= "spp") %>% drop_na() %>% group_by(classification, month) %>% mutate(total_biomass = sum(Biomass)) %>% #sum initial biomass of each species classification by month
ungroup() %>% mutate(percent_biomass= Biomass/total_biomass) %>% #divide initial biomass for a species by total biomass for a species group
mutate(weighted_biomass = biomass*percent_biomass) %>%
#get weighted biomass by multiplying relative biomass by percent biomass 
group_by(classification, month) %>% summarize(agg_rel_biomass= sum(weighted_biomass))
#get aggregated relative biomass for each species classification by summing weighted biomass for each species classification for each month  
mesopelagics_only <- data_2 %>% filter(spp== "mesopelagics") %>% rename(Mesopelagics= "spp")
ggplot() + geom_line(data=data_2_g1, aes(x=month, y=agg_rel_biomass, col= as.factor(classification))) + geom_line(data=mesopelagics_only, aes(x = month, y = biomass, col= "Mesopelagics")) + labs(col= "Species classification") +
  xlab("Month") + ylab("Relative biomass") + 
scale_color_discrete(type = "viridis") + standard_theme 

#Graph 2: predators of mesopelagic fish, mesopelagic fish, mesopelagic prey that predators of mesopelagic fish prey on

#Identify prey of mesopelagics where higher diet proportion predators of mesopelagic fish have at least 0.05 in their diet 
prey_of_mesopelagic_preds <- Diet[c(3,6:10,18), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% rename(spp= DIETS) %>% mutate(classification= "Prey of mesopelagics") %>% select(-value, -name)

combo_g2 <- rbind(prey_of_mesopelagic_preds, predators_mesopelagic_0.05)
combo_g2 <-combo_g2[!duplicated(combo_g2$spp),] 

GroupInfo_edt <- GroupInfo_CalCur_USE %>% select(Group, Biomass) %>% rename(spp= Group) 
data_2_g2 <- data_2 %>% left_join(GroupInfo_edt, by= "spp") %>% left_join(combo_g2, by= "spp") %>% drop_na() %>% group_by(classification, month) %>% mutate(total_biomass = sum(Biomass)) %>% 
ungroup() %>% mutate(percent_biomass= Biomass/total_biomass) %>% 
mutate(weighted_biomass = biomass*percent_biomass) %>%
group_by(classification, month) %>% summarize(agg_rel_biomass= sum(weighted_biomass))

mesopelagics_only <- data_2 %>% filter(spp== "mesopelagics") %>% rename(Mesopelagics= "spp")
ggplot() + geom_line(data=data_2_g2, aes(x=month, y=agg_rel_biomass, col= as.factor(classification))) + geom_line(data=mesopelagics_only, aes(x = month, y = biomass, col= "Mesopelagics")) + labs(col= "Species classification") +
  xlab("Month") + ylab("Relative biomass") + 
scale_color_discrete(type = "viridis") + standard_theme

#Graph 3: combination of graph 1 and 2
combo_g3 <- rbind(prey_of_mesopelagic_preds[!duplicated(prey_of_mesopelagic_preds$spp),], predators_mesopelagic_0.05, pred_mesopelagic_prey_0.05[!duplicated(pred_mesopelagic_prey_0.05$spp),])

data_2_g3 <- data_2 %>% left_join(GroupInfo_edt, by= "spp") %>% left_join(combo_g3, by= "spp") %>% drop_na() %>% group_by(classification, month) %>% mutate(total_biomass= sum(unique(Biomass))) %>% ungroup() %>% mutate(percent_biomass= Biomass/total_biomass) %>% mutate(weighted_biomass= biomass*percent_biomass) %>% group_by(classification,month) %>% summarize(agg_rel_biomass= sum(weighted_biomass))

mesopelagics_only <- data_2 %>% filter(spp== "mesopelagics") %>% rename(Mesopelagics= "spp")
ggplot() + geom_line(data=data_2_g3, aes(x=month, y=agg_rel_biomass, col= as.factor(classification))) + geom_line(data=mesopelagics_only, aes(x = month, y = biomass, col= "Mesopelagics")) + labs(col= "Species classification") + scale_x_continuous(name= "Year", breaks= c(0, 120, 240, 360, 480, 600), labels= c("2001", "2010", "2020", "2030", "2040", "2050")) +geom_hline(yintercept=1, linetype= "solid", color= "black", size= 0.75) +
  xlab("Month") + ylab("Relative biomass") + 
scale_color_discrete(type = "viridis") + standard_theme + ggsave(width = 6, height = 4, "/Users/sallydowd/Documents/OTZ.manuscript/Images/Sub.2/preds.otherprey.25.jpeg", dpi=600)
```

#Ungrouped graph: In this case the other prey is predators of mesopelagic fish that have >= 5% of their diet depend on prey of mesopelagic fish 
#Show non-intuitive predators (ones that increased) + show major prey of mesopelagics
```{r}
#Graph 1: two predators of mesopelagic fish (>= 0.05) that increased and two that decreased the most, mesopelagic fish, the prey of mesopelagic fish that these predators relied most on, the other prey that these predators relied most on 
pmp_df <- prey_of_mesopelagic_preds[!duplicated(prey_of_mesopelagic_preds$spp),] #prey of mesopelagic fish that make up at least 5% of diet for some predators of mesopelagic fish 
pmf_df <- predators_mesopelagic_0.05 #High diet proportion predators of mesopelagic fish that consume at least 5% of mesopelagic fish prey
combo_u1 <- rbind(pmp_df, pmf_df, mesopelagics_edt)

mesopelagics_edt <- mesopelagics_only %>% rename(spp= "Mesopelagics") %>% mutate(classification= "Mesopelagics") %>% select(spp, classification)

data_2_u1 <- data_2 %>% left_join(combo_u1, by= "spp") %>% drop_na()

data_2_u1 %>% filter(spp == c("Leach's S. Petrel", "copepods", "Dolphins", "Juv. hake", "Grenadiers", "euphausiids", "Other cephal.", "mesopelagics")) %>% ggplot() + geom_line(aes(x=month, y=biomass, col= spp, linetype= classification)) + scale_color_discrete(type= "viridis", labels= c("copepods"= "Copepods", "euphausiids"= "Euphausiids", "Leach's S. Petrel"= "Leach's s. petrel", "mesopelagics"= "Mesopelagics", "Other cephal." = "Other cephalopods")) + scale_x_continuous(name= "Year", breaks= c(0, 120, 240, 360, 480, 600), labels= c("2001", "2010", "2020", "2030", "2040", "2050")) +  geom_hline(yintercept=1, linetype= "solid", color= "black", size= 0.75) + xlab("Month") + ylab("Relative biomass") + labs(col= "Species", linetype= "Species classification") + standard_theme + guides(linetype = guide_legend(order = 2),col = guide_legend(order = 1)) + ggsave(width = 6, height = 4, "/Users/sallydowd/Documents/OTZ.manuscript/Images/Sub.2/mix.species.25.jpeg", dpi=600)


Diet[,c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% filter(name== "Grenadiers")


```

#Weighted biomass for diet proportions of mesopelagic predators, mesopelagic fish, and prey of mesopelagic fish (that predators of mesopelagic fish prey on for >= 5% of their diet)
```{r}
prey_of_mesopelagic_preds <- Diet[c(3,6:10,18), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% rename(spp= name) #prey of mesopelagics where higher diet proportion predators of mesopelagic fish have at least 

avg_pmp <- prey_of_mesopelagic_preds %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for predators on mesopelagic prey

avg_pmf <- predators_of_mesopelagic[15,] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for mesopelagic predators on mesopelagic fish

avg_mfmp <- Diet[c(3,6,7, 8:10, 18), c(1,16)] %>% summarize(sum(mesopelagics))
#non-weighted value (just sum) of diet composition for mesopelagic fish on mesopelagic prey because would be dividing Biomass of mesopelagics/total_biomass of mesopelagics which would equal 1 

avg_mpp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for mesopelagic prey on mesopelagic predators

avg_mfp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,16)] %>% summarize(mean(mesopelagics))
#weighted value of diet composition for mesopelagic fish on mesopelagic predators

avg_mpmf <- Diet[15, c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for mesopelagic prey on mesopelagic fish

avgsmpmp <- Diet[c(3,6,7, 8:10, 18), c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight)) 
 #weighted value of diet composition for mesopelagic prey on mesopelagic prey  
  
avg_pp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight)) 
#weighted value of diet composition for mesopelagic predators on mesopelagic predators
```

```{r}

                                                                                      
```


#Exploring why euphausiids and forage fish increase
```{r}
forage <- Cal_bal$Group[19:27]

plot_forage <- data_2 %>% filter(spp %in% forage) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= factor(spp))) + scale_color_discrete(type= "viridis", labels= c("Pacific Mackerel" = "Pacific mackerel")) + xlab("Month") + ylab("Relative biomass") +
 labs(colour= "Forage fish", title= "B") + standard_theme #+ ggsave(width = 6, height = 6, "~/Documents/OTZ.manuscript/Images/plot.noncommercial.50.jpeg")
plot_forage


```


