---
title: "Diet.description"
output: html_document
---

#Load Libraries
```{r}
library(dplyr)
library(readr)
library(Rpath)
library(tidyverse)

GroupInfo_CalCur_USE <- read_csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/GroupInfo_CalCur-USE.csv")

Diet <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Diet_CalCur_USE.csv")  
Diet <- Diet %>% rename("Juv. hake" = "Juv..hake", "Yellowtail rock."= "Yellowtail.rock.", "Shelf rock."= "Shelf.rock.", "P. Ocean Perch"= "P..Ocean.Perch", "Longspine thorny."= "Longspine.thorny.", "Tufted Puffin"= "Tufted.Puffin", "Leach's S. Petrel"= "Leach.s.S..Petrel","Fin whale"= "Fin.whale", "Slope rock."= "Slope.rock.", "carniv-zoops"= "carniv.zoops", "small jellies"= "small.jellies", "large jellies"= "large.jellies", "pandalid shp"= "pandalid.shp", "benthic shp"= "benthic.shp", "tanner crab"= "tanner.crb", "Benthic fish"= "Benthic.fish", "surf perch"= "surf.perch", "Other cephal."= "Other.cephal.", "Market squid"= "Market.squid", "Pacific Mackerel"= "Pacific.Mackerel")
Diet
```

#Functions needed- turn R markdown into R script eventually so can just source the functions
```{r}
target_sp <- "mesopelagics"
target_F <- 0.25
names(target_F) <- target_sp 

library(data.table)
scene2 <- copy(base_sim_scene)
run2 <- rsim.run(scene2, method = 'AB', years=2001)
for (yr in 2002:2050){
  # stock assessment based on current biomass * error
  # use this if including two species: assessment <- end_biomass(run3)[target_sp] * (runif(2) + 0.5)   
assessment <- end_biomass(run2)[target_sp] * (runif(length(target_sp)) + 0.5) #argument of length 0 means input is of length zero 
assessment
  # convert target F * assessment to a total catch 
  catches <- target_F * assessment
  # apply the catch to mesopelagic fish
  for (sp in target_sp){
    scene2 <- adjust.fishing(scene2, "ForcedCatch", sp, yr, value=catches[sp])
  }
  #run 1 year
  run2 <- rsim.step(scene2, run2, method = 'AB', year.end=yr)  
}  
rsim.plot(run2,all_species)  

rsim.plot.fun <- function(Rsim.output, spname, indplot = F, ...){
  opar <- par(no.readonly = T)
  if(indplot == F){
    biomass <- Rsim.output$out_Biomass[, 2:ncol(Rsim.output$out_Biomass)]
    n <- ncol(biomass)
    start.bio <- biomass[1, ]
    start.bio[which(start.bio == 0)] <- 1
    rel.bio <- matrix(NA, dim(biomass)[1], dim(biomass)[2])
    for(isp in 1:n) rel.bio[, isp] <- biomass[, isp] / start.bio[isp]
  }
  rel.bio
} 

names_spp <- run1$params$spname[2:94]

standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.text.align= 0, legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) 

```

#Define datasets
```{r}
predators_of_mesopelagic <- Diet[,c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]
mesopelagic_pred <- Diet[15, c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]
prey_of_mesopelagic <- Diet[c(3,6:10,15,18), c(1,16)]
forage.pred <- Diet[19:27, c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]
euphausiids <- Diet[7, c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64, 66:68, 70:71, 73:74, 77:81, 83, 86:89)]

species <- as.data.frame(forage.pred$DIETS) %>% mutate(X=1:9) %>% rename("Species"= "forage.pred$DIETS")
#Check if this is right 
forage.pred$MAX <- apply(forage.pred[,-1],1,max)
forage <- as.data.frame(forage.pred$MAX) %>% mutate(X=1:9)
forage <- forage %>% rename("Max diet"= "forage.pred$MAX") %>% left_join(species, by= "X") %>% select(-X)
forage
mesopelagic.pred$MAX <- apply(mesopelagic.pred[,-1],1,max,na.rm=TRUE)
mesopelagic <- as.data.frame(mesopelagic.pred$MAX) %>% rename("Max diet"= "mesopelagic.pred$MAX") %>% mutate("Species"= "mesopelagic fish")
euphausiids$MAX <- apply(euphausiids[,-1],1,max,na.rm=TRUE)
euphausiids <- as.data.frame(euphausiids$MAX) %>% rename("Max diet"= "euphausiids$MAX") %>% mutate("Species"= "euphausiids")

combination <- bind_rows(forage, mesopelagic, euphausiids)

predators_of_mesopelagic[15,]
```

#Predators of mesopelagic fish: diet dependence on mesopelagics and forage fish
```{r}
#For each predator of mesopelagic fish find their prey items that make up at least 5% of their diet
predators_of_mesopelagic

 #Webplot highlighting predator and prey connections to mesopelagic fish  
webplot(Cal_bal, eco.name= attr(Cal_bal, "California Current"), line.col="grey", labels=TRUE, highlight= "mesopelagics", highlight.col= c("blue", "red", "NA"), label.num=TRUE, label.cex=0.7, fleets=FALSE)
```

```{r}
#Make run2 a dataframe 
z_2 <-rsim.plot.fun(run2)
z_2<-as.data.frame(z_2)
names(z_2) <- names_spp
z_2$month <- 1:600
library(tidyr)
data_2 <- gather(z_2, key = spp, value = biomass, -month)

data_2 %>% filter(spp %in% c(predators_of_mesopelagic, mesopelagic_pred))

frate.table(scene0)
```

#Grouped graph: weighted aggregate biomass for predators of mesopelagic fish, other prey of predators of mesopelagic fish (not just prey of mesopelagic fish)
```{r}
#Pred_mesopelagic_prey_0.05 is made up of the predators of mesopelagic fish that have other prey consisting of at least 5% of their diet, predators_mesopelagic_0.05 are the predators of mesopelagic fish that have at least 5% of their diet consist of mesopelagics 
pred_mesopelagic_prey_0.05 <- predators_of_mesopelagic %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% select(DIETS) %>% unique() %>% rename(spp = DIETS) %>% mutate(classification= "Other prey") #Make dataset longer by increasing number of rows, only select prey when at least one predators depends on them for >= 5% of their diet
#Remove mesopelagics as predator of mesopelagics, not included as non-commercial or commercial predator in the study
predators_mesopelagic_0.05 <- predators_of_mesopelagic[15,] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% select(name) %>% mutate(classification= "Predators") %>% rename(spp= name)
combo <- rbind(pred_mesopelagic_prey_0.05, predators_mesopelagic_0.05)

GroupInfo_edt <- GroupInfo_CalCur_USE %>% select(Group, Biomass) %>% rename(spp= Group) #select original biomass from group info 
data_2_edt <- data_2 %>% left_join(GroupInfo_edt, by= "spp") %>% left_join(combo, by= "spp") %>% drop_na() %>% group_by(classification, month) %>% mutate(total_biomass = sum(Biomass)) %>% #sum initial biomass of each species classification by month
ungroup() %>% mutate(percent_biomass= Biomass/total_biomass) %>% #divide initial biomass for a species by total biomass for a species group
mutate(weighted_biomass = biomass*percent_biomass) %>%
#get weighted biomass by multiplying relative biomass by percent biomass 
group_by(classification, month) %>% summarize(agg_rel_biomass= sum(weighted_biomass))
#get aggregated relative biomass for each species classification by summing weighted biomass for each species classification for each month  
mesopelagics_only <- data_2 %>% filter(spp== "mesopelagics") %>% rename(Mesopelagics= "spp")
ggplot() + geom_line(data=data_2_edt, aes(x=month, y=agg_rel_biomass, col= as.factor(classification))) + geom_line(data=mesopelagics_only, aes(x = month, y = biomass, col= "Mesopelagics")) + labs(col= "Species classification") +
  xlab("Month") + ylab("Relative biomass") + 
scale_color_discrete(type = "viridis") + standard_theme + ggsave(width = 6, height = 6, "/Users/sallydowd/Documents/OTZ.manuscript/Images/preds.otherprey.25.jpeg")
```

#Ungrouped graph: In this case the other prey is predators of mesopelagic fish that have >= 5% of their diet depend on prey of mesopelagic fish 
#Show non-intuitive predators (ones that increased) + show major prey of mesopelagics
```{r}
pmp_df <- Diet[c(3,6:10,18), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% select(DIETS) %>% rename(spp= DIETS) %>% mutate(Classification= "Mesopelagic fish prey")
pmp_df <- prey_of_mesopelagic_preds[!duplicated(prey_of_mesopelagic_preds$spp),] #High diet proportion predators of mesopelagic fish that consume at least 5% of mesopelagic fish prey
pmf_df <- predators_of_mesopelagic[15,] %>% rename("Juv. hake" = "Juv..hake", "Yellowtail rock."= "Yellowtail.rock.", "Shelf rock."= "Shelf.rock.", "P. Ocean Perch"= "P..Ocean.Perch", "Longspine thorny."= "Longspine.thorny.", "Tufted Puffin"= "Tufted.Puffin", "Leach's S. Petrel"= "Leach.s.S..Petrel","Fin whale"= "Fin.whale", "Slope rock."= "Slope.rock.") %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% select(name) %>% mutate(Classification= "Predators") %>% rename(spp=name)
combo <- rbind(pmp_df, pmf_df)
data_2_edt <- data_2 %>% left_join(combo, by= "spp") %>% drop_na()

data_2_edt %>% 
filter(spp == c("Fin whale", "Yellowtail rock.", "Juv. hake", "Grenadiers", "euphausiids", "copepods", "carniv-zoops")) %>% ggplot() + geom_line(aes(x=month, y=biomass, col= spp, linetype= Classification))
  

predators_of_mesopelagic %>% pivot_longer(-DIETS) %>% filter(value >= 0.05, name== "Grenadiers")
#Fin whale: euphausiids as main, yellowtail. rock.: euphausiids as main, Juv. hake: copepods as main, other cephal. important for grenadiers

+ geom_line(data=mesopelagics_only, aes(x = month, y = biomass, col= "Mesopelagics")) + labs(col= "Species classification") +
  xlab("Month") + ylab("Relative biomass") + 
scale_color_discrete(type = "viridis") + standard_theme + ggsave(width = 6, height = 6, "/Users/sallydowd/Documents/OTZ.manuscript/Images/preds.otherprey.25.jpeg")
```

```{r}
#Identify prey of mesopelagics where higher diet proportion predators of mesopelagic fish have at least 0.05 in their diet 
prey_of_mesopelagic_preds <- Diet[c(3,6:10,18), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% rename(spp= name) 

data_2_edt <- data_2 %>% left_join(GroupInfo_edt, by= "spp") %>% left_join(combo, by= "spp") %>% drop_na() %>% group_by(classification, month) %>% mutate(total_biomass = sum(Biomass)) %>% #sum initial biomass of each species classification by month
ungroup() %>% mutate(percent_biomass= Biomass/total_biomass) %>% #divide initial biomass for a species by total biomass for a species group
mutate(weighted_biomass = biomass*percent_biomass) %>%
#get weighted biomass by multiplying relative biomass by percent biomass 
group_by(classification, month) %>% summarize(agg_rel_biomass= sum(weighted_biomass))
#get aggregated relative biomass for each species classification by summing weighted biomass for each species classification for each month  

Diet

avg_pmp <- prey_of_mesopelagic_preds %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(Biomass)) %>% mutate(percent_biomass= Biomass/total_biomass) %>% mutate(weighted_biomass = 1*percent_biomass) %>% mutate(weighted_diet= weighted_biomass*value) %>% summarize(mean(weighted_diet))


%>% summarize(mean(value)) #mean value of diet composition for predators on mesopelagic prey
avg_pmf <- predators_of_mesopelagic[15,] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% summarize(mean(value)) #mean value of diet composition for mesopelagic predators on mesopelagic fish
avg_mfmp <- Diet[c(3,6,7, 8:10, 18), c(1,16)] %>% summarize(mean(mesopelagics)) #mean value of diet composition for mesopelagic fish on mesopelagic prey
avg_mpp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% summarize(mean(value)) #mean value of diet composition for mesopelagic prey on mesopelagic predators
avg_mfp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,16)] %>% summarize(mean(value))
#mean value of diet composition for mesopelagic fish on mesopelagic predators
avg_mpmf <- Diet[15, c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% summarize(mean(value)) #mean value of diet composition for mesopelagic prey on mesopelagic fish
avgsmpmp <- Diet[c(3,6,7, 8:10, 18), c(1,4,7,8, 9:11, 19)]  %>% pivot_longer(-DIETS) %>% summarize(mean(value))
  #mean value of diet composition for mesopelagic prey on mesopelagic prey
avg_pp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% summarize(mean(value))
 #mean value of diet composition for mesopelagic predators on mesopelagic predators
```

#Exploring why euphausiids and forage fish increase
```{r}
forage <- Cal_bal$Group[19:27]

plot_forage <- data_2 %>% filter(spp %in% forage) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= factor(spp))) + scale_color_discrete(type= "viridis", labels= c("Pacific Mackerel" = "Pacific mackerel")) + xlab("Month") + ylab("Relative biomass") +
 labs(colour= "Forage fish", title= "B") + standard_theme #+ ggsave(width = 6, height = 6, "~/Documents/OTZ.manuscript/Images/plot.noncommercial.50.jpeg")
plot_forage


```


