---
title: "100 Rsim runs"
output: html_document
---

#Code to analyze approximately 500 datasets from the Monte Carlo analysis in Koehn et al. (2016)
```{r}
library(dplyr)
library(readr)
library(Rpath)
library(data.table)
library(tidyr)
library(ggplot2)
```

#Load this each time 
```{r}
GroupInfo_CalCur_USE <- read.csv("../Data/Groupinfo_CalCur-USE.csv")  
Diet_CalCur_USE <- read.csv("../Data/Diet_CalCur_USE.csv")  

## need to add row and column names from example files
base_list <-list.files(path = "../Data/500BaseParameters/",
               pattern = "*.csv", 
               full.names = T)
diet_list <-list.files(path = "../Data/DietMatrices/",
               pattern = "*.csv", 
               full.names = T)
Cbase_list <- list()
Cdiet_list <- list()
```

#Don't need to run this again! Was just to modify datasets originally (have already saved them)
```{r}
#Add columns to base parameters
#groups <- GroupInfo_CalCur_USE %>%
 # select(Group, Type, Unassim, BioAcc, ProdCons, DetInput, DetritusFate, Fleet1_Discards) %>%
  #mutate(X = 1:94)
#groups

#Extract row 94 from Group info (Fleet 1)
#Fleet1 <- GroupInfo_CalCur_USE[94, ] %>% select(Biomass, EE, PB, QB, Fleet1_Landings) %>% mutate(X= 94)

#Rename rows in diet files 
#groups2 <- Diet_CalCur_USE %>% select(DIETS) %>% mutate(X= 1:94)

#Rename columns in diet file
#group_vector <- groups2$DIETS[1:92]
#group_vector <- as.data.frame(c("DIETS", group_vector))
#group_vector <- groups2[,"DIETS"]

#test <- read_csv("../Data/DietMatrices/diet100227.csv")
#test_vector <- colnames(test)[1:93]

#First loop for group: rename columns, add Fleet 1 row (assigned position X=94), add new parameters (columns) onto dataset, redu column order
#for (i in 1:500) {
# Cbase_list[[i]] <- read.csv(base_list[i]) %>% 
    #rename(Biomass = "V1", PB = "V2", QB = "V3", EE = "V4", Fleet1_Landings = "V5")
 #Cbase_list[[i]] <- bind_rows(Fleet1, Cbase_list[[i]]) %>% arrange(X) %>% left_join(groups, by= "X") %>% select(-X) %>% relocate(Group)
#col_order <- c("Group", "Biomass", "EE", "PB", "QB", "Type", "Unassim", "BioAcc", "ProdCons", "DetInput", "DetritusFate", "Fleet1_Landings", "Fleet1_Discards")
#Cbase_list[[i]] <- Cbase_list[[i]][, col_order]
#write.csv(Cbase_list[[i]], paste("output", base_list[i]))
#}

#z<- read_csv(base_list[1])

#colnamesgroups<- c("DIETS", group_vector[1:93])
#Diet matrices: #REMOVE IMPORT ROW??? ROW 94 (V93) IN TEST DATA
#for (i in 1:500) {
#Cdiet_list[[i]] <- read_csv(diet_list[i], col_names = colnamesgroups ) %>% select(-Detritus)
#Cdiet_list[[i]]<- Cdiet_list[[i]][-1,]
#Cdiet_list[[i]] <- Cdiet_list[[i]] %>%
 # mutate(DIETS = c(colnamesgroups[2:94], "Import"))
#write.csv(Cdiet_list[[i]], diet_list[i])
#}

#z2 <- read_csv(diet_list[1])

```

#Need to load in some data/functions from OTZ.results.newRpath.Rmd file to make some of the data manipulation easier
```{r}
#Original scenario
Cbase <- "../Data/Groupinfo_CalCur-USE.csv"  
Cdiet <- "../Data/Diet_CalCur_USE.csv"  
#Load unbalanced Ecopath model
CalCur_unbal <- read.rpath.params(Cbase, Cdiet)
check.rpath.params(CalCur_unbal)
#Balance original Ecopath model
Cal_bal <- rpath(CalCur_unbal) 
Cal_bal

#Base Scenario
yearlist = 2001:2050
base_sim_scene <- rsim.scenario(Cal_bal, CalCur_unbal, years = yearlist)
scene0 <- base_sim_scene
run0 <- rsim.run(scene0, method='AB', years=yearlist) 

#End bimoass function 
end_biomass <- function(rsim){return(rsim$out_Biomass[dim(rsim$out_Biomass)[1], 2:(dim(rsim$out_Biomass)[2])])}
end_biomass

#Mesopelagic harvest level of 25% applied to original model
target_sp <- "mesopelagics"
target_F <- 0.25
names(target_F) <- target_sp 
scene2 <- copy(base_sim_scene)
run2 <- rsim.run(scene2, method = 'AB', years=2001)
for (yr in 2002:2050){
  # stock assessment based on current biomass * error
  # use this if including two species: assessment <- end_biomass(run3)[target_sp] * (runif(2) + 0.5)   
assessment <- end_biomass(run2)[target_sp] * (runif(length(target_sp)) + 0.5) #argument of length 0 means input is of length zero 
assessment
  # convert target F * assessment to a total catch 
  catches <- target_F * assessment
  # apply the catch to mesopelagic fish
  for (sp in target_sp){
    scene2 <- adjust.fishing(scene2, "ForcedCatch", sp, yr, value=catches[sp])
  }
  #run 1 year
  run2 <- rsim.step(scene2, run2, method = 'AB', year.end=yr)  
}  

#Original groups
noncommerical_predators <- c(Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[43:47], Cal_bal$Group[50:56], Cal_bal$Group[58:60], Cal_bal$Group[63], Cal_bal$Group[65:67], Cal_bal$Group[69:70], Cal_bal$Group[72:73], Cal_bal$Group[76:80], Cal_bal$Group[82], Cal_bal$Group[85:88])
noncommerical_predators
commercial_predators <- c(Cal_bal$Group[18], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[42:48], Cal_bal$Group[50:56], Cal_bal$Group[58:59])
commercial_predators
```

#Loop through 500 data sets!
#Reading in data, unbalanced model
```{r}
CalCur_unbal2_list <- list() #Creating an empty list to store vectors so can index 

  for (i in 1:500) {
  Cbase2 <-  base_list[[i]] 
  Cdiet2 <- diet_list[[i]]

#Edit the datasets
  CalCur_unbal2_list[[i]] <- read.rpath.params(Cbase2, Cdiet2)
  CalCur_unbal2_list[[i]][[1]] <- CalCur_unbal2_list[[i]][[1]] %>%
    select(-X)
  CalCur_unbal2_list[[i]][[1]][93, 5] <- NA
  CalCur_unbal2_list[[i]][[1]][CalCur_unbal2_list[[i]][[1]] == -1] <- NA
   CalCur_unbal2_list[[i]][[2]] <- CalCur_unbal2_list[[i]][[2]] %>%
    select(-X)

  check.rpath.params(CalCur_unbal2_list[[i]])
  }
```

#Creating 500 balanced Rpath models
```{r}
Cal_bal2 <- list()
for (i in 1:500) {
Cal_bal2[[i]] <- rpath(CalCur_unbal2_list[[i]]) 
  cbind(Cal_bal2[[i]]$Group, Cal_bal2[[i]]$EE)
  cbind(Cal_bal2[[i]]$Group, Cal_bal2[[i]]$TL) 
}
```


#Creating 500 different base scenarios 
```{r}
#Define species
all_species <- Cal_bal$Group[1:92] 

#Base scenario
base_sim_scene2 <- list()
scene0.2 <- list()
run0.2 <- list()
for (i in 1:500) {
yearlist = 2001:2050

base_sim_scene2[[i]] <- rsim.scenario(Cal_bal2[[i]], CalCur_unbal2_list[[i]], years = yearlist)
scene0.2[[i]] <- base_sim_scene2[[i]]
run0.2[[i]] <- rsim.run(scene0.2[[i]], method='AB', years=yearlist) 
  }

```

#Scenario 8: 25% harvest rate on mesopelagics for the 500 datasets
```{r}
target_sp <- "mesopelagics"
target_F <- 0.25
names(target_F) <- target_sp 

scene6.2 <- list()
run6.2 <- list()
assessment2 <- list()
catches2 <- list()
for (i in 1:500){
scene6.2[[i]] <- copy(base_sim_scene2[[i]])
run6.2[[i]] <- rsim.run(scene6.2[[i]], method = 'AB', years=2001)
for (yr in 2002:2050){
  # stock assessment based on current biomass * error
  # use this if including two species: assessment <- end_biomass(run3)[target_sp] * (runif(2) + 0.5)   
assessment2[[i]] <- end_biomass(run6.2[[i]])[target_sp] * (runif(length(target_sp)) + 0.5) #argument of length 0 means input is of length zero 
assessment2[[i]]
  # convert target F * assessment to a total catch 
  catches2[[i]] <- target_F * assessment2[[i]]
  # apply each catch to each species
  for (sp in target_sp){
    scene6.2[[i]] <- adjust.fishing(scene6.2[[i]], "ForcedCatch", sp, yr, value=catches2[[i]][sp])
  }
  # run 1 year
  run6.2[[i]] <- rsim.step(scene6.2[[i]], run6.2[[i]], method = 'AB', year.end=yr)  #Says in Rsim document on NOAA github that older versions of Ecosim use AB and new versions use 'RK4', default for Rpath is RK4 but we aren't using it??
}  
  }
rsim.plot(run6.2[[500]], all_species)
```

#Manipulating output data
```{r}
rsim.plot.fun <- function(Rsim.output, spname, indplot = F, ...){
  opar <- par(no.readonly = T)
  if(indplot == F){
    biomass <- Rsim.output$out_Biomass[, 2:ncol(Rsim.output$out_Biomass)]
    n <- ncol(biomass)
    start.bio <- biomass[1, ]
    start.bio[which(start.bio == 0)] <- 1
    rel.bio <- matrix(NA, dim(biomass)[1], dim(biomass)[2])
    for(isp in 1:n) rel.bio[, isp] <- biomass[, isp] / start.bio[isp]
  }
  rel.bio
} 

#Convert 500 rsim models into a data frame 
names_spp.2 <- run2$params$spname[2:94]
z_6.2 <- list()
data_6.2 <- list()
for (i in 1:500) {
z_6.2[[i]] <- rsim.plot.fun(run6.2[[i]])
z_6.2[[i]] <- as.data.frame(z_6.2[[i]])
names(z_6.2[[i]]) <- names_spp.2
z_6.2[[i]]$month <- 1:600
data_6.2[[i]] <- gather(z_6.2[[i]], key = spp, value = biomass, -month)
data_6.2[[i]] <- data_6.2[[i]] %>% filter(month==600) %>% mutate(X= i)
}

#Combine datasets into one: remove the 7 rsim simulations that are outliers
entire_data_6 <- bind_rows(data_6.2) %>% filter(X == 67)
entire_data_6 <- bind_rows(data_6.2) %>% filter(!X %in% c(67, 126, 237, 343, 396, 430, 472)) %>% arrange(spp)
entire_data_6 <- transform(entire_data_6,ID=as.numeric(factor(spp))) #Add column for ID that correlates to each different species 

write.csv(entire_data_6, "/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/entire.data.6.csv")

```

#Confidence intervals: Calculate 95% CI for each model at month 600 (last month in model) for commercially and non-commerically valuable species 

```{r}
read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/entire.data.6.csv")

#Calculate CI:
n <- 500
t_star <- 1.960
sample_mean <- list()
sd_error <- list()
CI_lower <- list()
CI_upper <- list()

for (i in unique(entire_data_6$ID)){
subdf <- subset(x=entire_data_6, subset=ID==i)
sample_mean[[i]] <- mean(subdf$biomass)
sd_error[[i]] <- (sd(subdf$biomass))/(sqrt(n))
CI_lower[[i]] <- sample_mean[[i]] - (t_star*sd_error[[i]])
CI_upper[[i]] <- sample_mean[[i]] + (t_star*sd_error[[i]])
}

df2 <- as.data.frame(do.call(rbind, CI_upper)) %>% rename("CI_upper" = V1) %>% mutate(ID = 1:93)
df1 <- as.data.frame(do.call(rbind, CI_lower)) %>% rename("CI_lower"= V1) %>% mutate(ID = 1:93) %>% left_join(df2, by = "ID") %>% relocate(ID)

#Calculate quantiles
quantile_0.025 <- list()
quantile_0.975 <- list()

for (i in unique(entire_data_6$ID)){
subdf2 <- subset(x=entire_data_6, subset=ID==i)
quantile_0.025[[i]] <- quantile(subdf2$biomass, 0.025)
quantile_0.975[[i]] <- quantile(subdf2$biomass, 0.975)
}

df_quant2 <- as.data.frame(do.call(rbind, quantile_0.975)) %>% mutate(ID = 1:93)
df_quant1 <- as.data.frame(do.call(rbind, quantile_0.025)) %>% mutate(ID = 1:93) %>% left_join(df_quant2, by = "ID") %>% relocate("ID")

sum_data_6 <- entire_data_6 %>% group_by(spp, ID) %>% summarise(mean_biomass = mean(biomass), sd_biomass= sd(biomass), median_biomass = median(biomass)) %>% left_join(df1, by = "ID")
sum_data_6 <- sum_data_6 %>% left_join(df_quant1, by = "ID") %>% rename(quant_2.5 = "2.5%", quant_97.5 = "97.5%")
sum_data_6 <- sum_data_6 %>% mutate("median.value" = cut(median_biomass, 
                      breaks = c(-Inf, 1, Inf),
                      labels = c("Decreasing", "Increasing"),
                      right = TRUE), "difference"= quant_97.5-quant_2.5)

write.csv(sum_data_6, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/sum_data_6.csv")
```

#Check to see if the seven datasets that produce unrealistic biomass estimates are outliers 
```{r}
rest_of_models_outliers <- entire_data_6 %>% group_by(spp) %>% summarize(IQR_biomass = IQR(biomass), quantile_lower = quantile(biomass, 0.25), quantile_upper = quantile(biomass, 0.75)) %>% mutate(lower_outlier = quantile_lower - (1.5*IQR_biomass), upper_outlier= quantile_upper + (1.5*IQR_biomass)) 

seven_models_outliers <- bind_rows(data_6.2) %>% filter(X %in% c(67, 126, 237, 343, 396, 430, 472)) %>% arrange(X)
seven_models_outliers <- transform(seven_models_outliers, ID.X=as.numeric(factor(X))) 
seven_models_outliers <- seven_models_outliers %>% spread(ID.X, biomass) %>% rename("model1"= "1", "model2"= "2", "model3"= "3", "model4"= "4", "model5"= "5", "model6"= "6", "model7" = "7")

#Each species final biomass in each of the 7 models are outliers 
test <- seven_models_outliers %>% left_join(rest_of_models_outliers) %>% mutate(outlier1 = ifelse(model1 < lower_outlier|model1 > upper_outlier, TRUE, FALSE), outlier2 = ifelse(model2 < lower_outlier|model2 > upper_outlier, TRUE, FALSE), outlier3 = ifelse(model3 < lower_outlier|model3 > upper_outlier, TRUE, FALSE), outlier4 = ifelse(model4 < lower_outlier|model4 > upper_outlier, TRUE, FALSE), outlier5 = ifelse(model5 < lower_outlier|model5 > upper_outlier, TRUE, FALSE), outlier6 = ifelse(model6 < lower_outlier|model6 > upper_outlier, TRUE, FALSE), outlier7 = ifelse(model7 < lower_outlier|model7 > upper_outlier, TRUE, FALSE))

```

#Graphs
```{r}
#Filter datasets for commercial and non-commercial predators of mesopelagic fish
new_noncomm <- noncommerical_predators[- c(9,7,11,4,6,17,1,12,8,10,19,5,16,18,15,14,13,3)] #Define group that has non-commercial predators that only had non-commercial value (not commercial value as well)
comm_value <- sum_data_6 %>% filter(spp %in% commercial_predators) 
write.csv(comm_value, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/comm_value.csv")
noncomm_value <- sum_data_6 %>% filter(spp %in% new_noncomm)
write.csv(noncomm_value, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/noncomm_value.csv")

#Plot CIs for commercial and non-commercial predators 
comm_value_500 <- comm_value %>% ggplot(aes(fill = median.value)) + 
  geom_pointrange(aes(y=spp, x=median_biomass, xmin=quant_2.5, xmax= quant_97.5), size=0.5, shape=22) + 
  scale_fill_grey() + 
scale_y_discrete(limits = unique(rev(comm_value$spp)), labels= c("Greenstriped" = expression(bold("Greenstriped rockfish")), "P. Ocean Perch"= expression(bold("P. ocean perch")), "Albacore" = expression(bold("Albacore")), "Arrowtooth" = expression(bold("Arrowtooth flounder")), "Canary rock."= expression(bold("Canary rockfish")), "Lingcod"= expression(bold("Lingcod")), "Longspine thorny."= expression(bold("Longspine thornyhead")), "Other cephal."= expression(bold("Other cephalopod")), "Sablefish"= expression(bold("Sablefish")), "Salmon"= expression(bold("Salmon")), "Sharks"= expression(bold("Sharks")), "Shelf rock."= expression(bold("Shelf rockfish")), "Shortspine thorny."= expression(bold("Shortspine thornyhead")), "Skates"= expression(bold("Skates")), "Slope rock."= expression(bold("Slope rockfish")), "Splitnose rock."= expression(bold("Splitnose rockfish")), "Widow rock."= expression(bold("Widow rockfish")), "Yellowtail rock."= expression(bold("Yellowtail rockfish")))) +
theme(legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10, family= "Helvetica
serif"), axis.title=element_text(size=12)) + 
geom_vline(xintercept = 1, size = 0.5) + 
xlab("Relative biomass") + ylab("Species") + labs(fill = "Median biomass", title= "A") + 
theme_bw() + theme(panel.border = element_blank()) +
theme(axis.line = element_line(colour = "black")) +
theme(legend.position= "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
ggsave(width = 6, height = 4, "~/Documents/OTZ.Manuscript/Images/500.commercial.25.jpeg")
comm_value_500

noncomm_value_500 <- noncomm_value %>% ggplot(aes(fill = median.value)) + geom_pointrange(aes(y=spp, x=median_biomass, xmin=quant_2.5, xmax= quant_97.5), size=0.5, color="black", shape=22) + 
scale_fill_grey() + 
scale_y_discrete(limits = unique(rev(noncomm_value$spp)), labels= c("Western Gull" = "Western gull", "Tufted Puffin" = "Tufted puffin", "Sea Lions"= "Sea lions", "Resident Orcas"= "Resident orcas", "Leach's S. Petrel" ="Leach's s. petrel", "Juv. Ele. Seal"="Juv. elephant seal", "Double corm."= "Double cormorant", "Brandt's corm." = "Brandt's cormorant", "Juv. rock." = "Juv. rockfish")) + 
geom_vline(xintercept = 1, size = 0.5) + 
xlab("Relative biomass") + ylab("Species") + labs(fill = "Median biomass", title= "B") + 
theme_bw() + theme(panel.border = element_blank()) +
theme(axis.line = element_line(colour = "black")) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
theme(legend.position= "none", legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) + 
ggsave(width = 6, height = 4, "~/Documents/OTZ.Manuscript/Images/500.noncommercial.25.jpeg")
noncomm_value_500
```

#Graphs
```{r}
#High diet proportion
sum_2_1 <- sum_data_6 %>% filter(spp %in% diet_over_0.05) %>% rename(species= "spp") %>% left_join(Predator_value, by= "species")
sum_2_1[5,12] <- "Neither"
sum_2_1$Value <- factor(sum_2_1$Value, levels = c("Commercial", "Non-commercial", "Both", "Neither"))
sum_2_1_graph <- sum_2_1 %>% ggplot(aes(col= Value)) + 
  geom_pointrange(aes(y=reorder(species, desc(species)), x=median_biomass, xmin=quant_2.5, xmax= quant_97.5), size=0.5, shape=22, fill= "white") + 
  scale_color_manual(values = c("#FF0000","#0000FF", "#00FF70", "#FF00FF")) + 
scale_y_discrete(labels= c("Tufted Puffin"= "Tufted puffin", "P. Ocean Perch"= "P. ocean perch", "Leach's S. Petrel"= "Leach's s. petrel", "Longspine thorny."= "Longspine thornyhead", "Shelf rock."= "Shelf rockfish", "Slope rock."= "Slope rockfish", "Yellowtail rock."= "Yellowtail rockfish")) + 
geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) +
  xlim(0.70, 1.35) + xlab("Relative biomass") + ylab("Species with high diet proportion") + labs(col = "Value", title= "G") + standard_theme + theme(axis.text.y=element_blank()) + theme(legend.position = "none") +
ggsave(width = 8, height = 7, "~/Documents/OTZ.Manuscript/Images/Sub.2/500.dietover.25.jpeg")
sum_2_1_graph

sum_2_2 <- sum_data_6 %>% filter(spp %in% diet_under_0.05) %>% rename(species= "spp") %>% left_join(Predator_value, by= "species")
#Remove mesopelagics
sum_2_2 <- sum_2_2[-15,]
sum_2_2$Value <- factor(sum_2_2$Value, levels = c("Non-commercial", "Both", "Neither"))
sum_2_2_graph <- sum_2_2 %>% ggplot(aes(col= Value)) + 
  geom_pointrange(aes(y=reorder(species, desc(species)), x=median_biomass, xmin=quant_2.5, xmax= quant_97.5), size=0.5, shape=22, fill= "white") + 
  scale_color_manual(values = c("#0000FF", "#00FF70")) +
 scale_y_discrete(labels= c("Western Gull"= "Western gull", "Resident Orcas"= "Resident orcas", "Juv. Ele. Seal"= "Juv. n. elephant seal", "Greenstriped"= "Greenstriped rockfish", "Double corm."= "Double cormorant", "Brandt's corm."= "Brandt's cormorant", "Arrowtooth"= "Arrowtooth flounder", "Canary rock."= "Canary rockfish", "Juv. rock."= "Juv. rockfish", "Other cephal."= "Other cephalopod", "Shortspine thorny."= "Shortspine thornyhead", "Widow rock."= "Widow rockfish", "mesopelagics"= "Mesopelagics", "Sea Lions"= "Sea lions", "Splitnose rock."= "Splitnose rockfish")) +
  geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) +
  xlim(0.70, 1.35) + xlab("Relative biomass") + ylab("Species with low diet proportion") + labs(col = "Value", title= "H") + standard_theme + 
ggsave(width = 8, height = 7, "~/Documents/OTZ.Manuscript/Images/Sub.2/500.dietunder.25.jpeg")
sum_2_2_graph

commercial_predators_df <- as.data.frame(commercial_predators) %>% rename(species= commercial_predators)
noncommercial_predators_df <- as.data.frame(noncommerical_predators) %>% rename(species= noncommerical_predators)
combo <- rbind(commercial_predators_df, noncommercial_predators_df, by= "species") %>% mutate(Dependence= 1)
sum_no_dependence <- sum_data_6 %>% rename(species= "spp") %>% left_join(combo, by= "species")
sum_no_dependence[is.na(sum_no_dependence)] <- 0 #Functional groups with no dependence on mesopelagics have a 0
sum_no_dependence <-sum_no_dependence[!duplicated(sum_no_dependence$species),] 
sum_no_dependence <- sum_no_dependence[-c(42,52),] #Remove mesopelagics and juvenile hake- dependence on mesopelagics but don't have commercial or non-commercial value so will be weird with filtering
sum_no_dependence_final <- sum_no_dependence %>% filter(Dependence==0)
#4: No dependence on mesopelagics
sum_2_3 <- sum_no_dependence_final
sum_2_3_graph <- sum_2_3 %>% ggplot() + 
  geom_pointrange(aes(y=reorder(species, desc(species)), x=median_biomass, xmin=quant_2.5, xmax= quant_97.5), size=0.5, shape=22, fill= "white") + 
scale_y_discrete(labels= c("Adult N. Ele. Seal"= "Adult n. elephant seal", "amphipods"= "Amphipods", "benthic shp"= "Benthic shrimp", "Brown Pelican"= "Brown pelican", "Cali. gull"= "California gull", "carniv-zoops" = "Carnivorous zooplankton", "copepods"= "Copepods", "Dungeness"= "Dungeness crab", "epibenthic"= "Epibenthic", "euphausiids"= "Euphausiids", "infauna"= "Infauna", "Juv. round."= "Juv. roundfish", "Juv. thorny."= "Juv. thornyhead", "large jellies"= "Large jellies", "micro-zoop"= "Microzooplankton", "Nearshore rock."= "Nearshore rockfish", "Pacific Mackerel"= "Pacific mackerel", "pandalid shp"= "Pandalid shrimp", "phytoplankton"= "Phytoplankton", "Pigeon Guillemot"= "Pigeon guillemot", "Rhino. auk"= "Rhinoceros auklet", "small jellies"= "Small jellies", "surf perch"= "Surf perch", "tanner crb"= "Tanner crab", "Transient Orcas"= "Transient orcas", "Yelloweye rock."= "Yelloweye rockfish")) +
  geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) +
  xlim(0.70, 1.35) + xlab("Relative biomass") + ylab("Species with low diet proportion") + labs(col = "Value", title= "I") + standard_theme + 
ggsave(width = 8, height = 7, "~/Documents/OTZ.Manuscript/Images/Sub.2/500.nodiet.25.jpeg")
sum_2_3_graph

library(ggpubr)

ggarrange(change_2_1_plot, change_1_1_plot, sum_2_1_graph  + rremove("y.text") + rremove("x.text"), 
          ncol = 3, nrow = 1) + 
ggsave(width = 12, height = 7, "~/Documents/OTZ.Manuscript/Images/Sub.2/proportion.combo.jpeg")
```

#Find the average of the input paramters (Biomass, QB, PB, and EE) from the seven outleir models and the 493 remaining models, trying to determine cause of unrealistic biomass estimates in the seven models
```{r}
#Biomass
seven_models <- CalCur_unbal2_list[c(67, 126, 237, 343, 396, 430, 472)]
rest_of_models <- CalCur_unbal2_list[-c(67, 126, 237, 343, 396, 430, 472)]

for (i in 1:493) {
rest_of_models[[i]] <- rest_of_models[[i]]$model %>% dplyr::select(Biomass, Group) 
}
seven_out <- list()
for (i in 1:7) {
seven_out[[i]] <- seven_models[[i]]$model %>% dplyr::select(Biomass, Group) 
}

Biomass_included <- bind_rows(rest_of_models) %>% group_by(Group) %>% summarize(mean_biomass = mean(Biomass)) %>% mutate(Model = "Included")
Biomass_excluded <- bind_rows(seven_out) %>% group_by(Group) %>% summarize(mean_biomass = mean(Biomass)) %>% mutate(Model = "Excluded")
Biomass_comp <- Biomass_included %>% bind_rows(Biomass_excluded) %>% arrange(Group)

#QB
seven_models <- CalCur_unbal2_list[c(67, 126, 237, 343, 396, 430, 472)]
rest_of_models <- CalCur_unbal2_list[-c(67, 126, 237, 343, 396, 430, 472)]
for (i in 1:493) {
rest_of_models[[i]] <- rest_of_models[[i]]$model %>% dplyr::select(QB, Group) 
}
seven_out <- list()
for (i in 1:7) {
seven_out[[i]] <- seven_models[[i]]$model %>% dplyr::select(QB, Group) 
}

QB_included <- bind_rows(rest_of_models) %>% group_by(Group) %>% summarize(mean_QB = mean(QB)) %>% mutate(Model = "Included")
QB_excluded <- bind_rows(seven_out) %>% group_by(Group) %>% summarize(mean_QB = mean(QB)) %>% mutate(Model = "Excluded")
QB_comp <- QB_included %>% bind_rows(QB_excluded) %>% arrange(Group)

#PB
seven_models <- CalCur_unbal2_list[c(67, 126, 237, 343, 396, 430, 472)]
rest_of_models <- CalCur_unbal2_list[-c(67, 126, 237, 343, 396, 430, 472)]
for (i in 1:493) {
rest_of_models[[i]] <- rest_of_models[[i]]$model %>% dplyr::select(PB, Group) 
}
seven_out <- list()
for (i in 1:7) {
seven_out[[i]] <- seven_models[[i]]$model %>% dplyr::select(PB, Group) 
}

PB_included <- bind_rows(rest_of_models) %>% group_by(Group) %>% summarize(mean_PB = mean(PB)) %>% mutate(Model = "Included")
PB_excluded <- bind_rows(seven_out) %>% group_by(Group) %>% summarize(mean_PB = mean(PB)) %>% mutate(Model = "Excluded")
PB_comp <- PB_included %>% bind_rows(PB_excluded) %>% arrange(Group)

#EE
seven_models <- CalCur_unbal2_list[c(67, 126, 237, 343, 396, 430, 472)]
rest_of_models <- CalCur_unbal2_list[-c(67, 126, 237, 343, 396, 430, 472)]
for (i in 1:493) {
rest_of_models[[i]] <- rest_of_models[[i]]$model %>% dplyr::select(EE, Group) 
}
seven_out <- list()
for (i in 1:7) {
seven_out[[i]] <- seven_models[[i]]$model %>% dplyr::select(EE, Group) 
}

EE_included <- bind_rows(rest_of_models) %>% group_by(Group) %>% summarize(mean_EE = mean(EE)) %>% mutate(Model = "Included")
EE_excluded <- bind_rows(seven_out) %>% group_by(Group) %>% summarize(mean_EE = mean(EE)) %>% mutate(Model = "Excluded")
EE_comp <- EE_included %>% bind_rows(EE_excluded) %>% arrange(Group)

#write.csv(Biomass_comp, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/500.inputs/Biomass_comp.csv")
#write.csv(QB_comp, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/500.inputs/QB_comp.csv")
#write.csv(EE_comp, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/500.inputs/EE_comp.csv")
#write.csv(PB_comp, "/Users/sallydowd/Documents/OTZ.manuscript/Excel/500.inputs/PB_comp.csv")
```
