---
title: "Untitled"
output: html_document
---

#Load required packages
```{r}
library(tidyverse)
library(data.table)
library(viridisLite)
library(viridis)
library(dplyr)
library(readxl)
library(ggplot2)
library(Rpath)
library(cowplot)
library(grid)
library(gridExtra)
```

#Importing code
```{r}
#Define code source for base and diet parameters
Cbase <- "../Data/Groupinfo_CalCur-USE.csv"  # Base biomass, production, fishing, etc.
Cdiet <- "../Data/Diet_CalCur_USE.csv"  # Diet matrix

#Load unbalanced Ecopath model from csv files
CalCur_unbal <- read.rpath.params(Cbase, Cdiet)

#Read in created parameter files with read.rpath.params function
check.rpath.params(CalCur_unbal)
```

#Running Ecopath in R: Balance an unbalanced Ecopath object
```{r}
#Balance the scenario using rpath function 
Cal_bal <- rpath(CalCur_unbal) 
Cal_bal

#Combine two data frames into a single data frame 
cbind(Cal_bal$Group, Cal_bal$EE)
cbind(Cal_bal$Group, Cal_bal$TL) 
```

#Define species for plotting (go over this)
```{r}
#Koehn's defined functional groups 
all_species <- Cal_bal$Group[1:92] 
mammals <- Cal_bal$Group[71:85] 
seabirds <- c(Cal_bal$Group[60:70], Cal_bal$Group[86:92])
pred_fish  <- Cal_bal$Group[33:59] 

#My customized defined functional groups 

#Plankton
Plankton <- c(Cal_bal$Group[1], Cal_bal$Group[5:10])
#Forage fish 
forage <- Cal_bal$Group[19:27]

#All predators of mesopelagic fish 
predators_mesopelagic <- c(Cal_bal$Group[15], Cal_bal$Group[18], Cal_bal$Group[28:29], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[42:48], Cal_bal$Group[50:56], Cal_bal$Group[58:60], Cal_bal$Group[63], Cal_bal$Group[65:67], Cal_bal$Group[69:70], Cal_bal$Group[72:73], Cal_bal$Group[76:80], Cal_bal$Group[82], Cal_bal$Group[85:88])
predators_mesopelagic

#Commercially valuable predators of mesopelagic fish
commercial_predators <- c(Cal_bal$Group[18], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[42:48], Cal_bal$Group[50:56], Cal_bal$Group[58:59])
commercial_predators

#Noncommercially valuable predators of mesopelagic fish
noncommerical_predators <- c(Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[43:47], Cal_bal$Group[50:56], Cal_bal$Group[58:60], Cal_bal$Group[63], Cal_bal$Group[65:67], Cal_bal$Group[69:70], Cal_bal$Group[72:73], Cal_bal$Group[76:80], Cal_bal$Group[82], Cal_bal$Group[85:88])
noncommerical_predators

#Diet proportion of mesopelagic fish less than 0.05
diet_under_0.05 <- c(Cal_bal$Group[15], Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[37], Cal_bal$Group[43:47], Cal_bal$Group[50], Cal_bal$Group[52:53], Cal_bal$Group[55], Cal_bal$Group[58:59], Cal_bal$Group[60], Cal_bal$Group[65:67], Cal_bal$Group[69],Cal_bal$Group[72:73], Cal_bal$Group[77:80], Cal_bal$Group[82], Cal_bal$Group[86:88])

diet_under_0.05_2 <- c(Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[37], Cal_bal$Group[43:47], Cal_bal$Group[50], Cal_bal$Group[52:53], Cal_bal$Group[55], Cal_bal$Group[58:59], Cal_bal$Group[60], Cal_bal$Group[65:67], Cal_bal$Group[69],Cal_bal$Group[72:73], Cal_bal$Group[77:80], Cal_bal$Group[82], Cal_bal$Group[86:88])
diet_under_0.05_2

#Diet proportion of mesopelagic fish equal to or greater than 0.05
diet_over_0.05 <- c(Cal_bal$Group[29], Cal_bal$Group[33], Cal_bal$Group[38], Cal_bal$Group[42], Cal_bal$Group[48], Cal_bal$Group[51], Cal_bal$Group[54], Cal_bal$Group[56], Cal_bal$Group[63], Cal_bal$Group[70], Cal_bal$Group[76], Cal_bal$Group[85])
diet_over_0.05

#Highest diet proportion 
highest_diet <- c(Cal_bal$Group[70], Cal_bal$Group[56], Cal_bal$Group[12])

#Predators of mesopelagic fish with annual economic value over $1/lb 
high.price <- c(Cal_bal$Group[47], Cal_bal$Group[45], Cal_bal$Group[38], Cal_bal$Group[55], Cal_bal$Group[46], Cal_bal$Group[43])
high.price
```

#Initialize a base scenario
```{r}
yearlist = 2001:2050

base_sim_scene <- rsim.scenario(Cal_bal, CalCur_unbal, years = yearlist)
scene0 <- base_sim_scene
run0 <- rsim.run(scene0, method='AB', years=yearlist) 
rsim.plot(run0,all_species)
```


#Helper function to look up the ending biomass of a scenario
```{r}
end_biomass <- function(rsim){return(rsim$out_Biomass[dim(rsim$out_Biomass)[1], 2:(dim(rsim$out_Biomass)[2])])}
end_biomass
```

#DEPLETION LEVEL OF 50%
#Run harvest scenario
```{r}
target_sp <- "mesopelagics" #Define target species
target_F <- 0.50 #Define target harvest rate 
names(target_F) <- target_sp 

scene1 <- copy(base_sim_scene)
run1 <- rsim.run(scene1, method = 'AB', years=2001)
for (yr in 2002:2050){
#Perform a stock assessment based on current biomass*error
assessment <- end_biomass(run1)[target_sp] * (runif(length(target_sp)) + 0.5)
assessment
#Calculate total catch by multiplying target harvest rate by assessment 
  catches <- target_F * assessment
  
#Apply each catch to the target species
  for (sp in target_sp){
    scene1 <- adjust.fishing(scene1, "ForcedCatch", sp, yr, value=catches[sp])
  }
  # run 1 year
  run1 <- rsim.step(scene1, run1, method='AB', year.end=yr) 
}  
rsim.plot(run1,all_species)  
```

#Plot biomass changes
```{r}
#Modify dataset to plot with ggplot2
rsim.plot.fun <- function(Rsim.output, spname, indplot = F, ...){
  opar <- par(no.readonly = T)
  if(indplot == F){
    biomass <- Rsim.output$out_Biomass[, 2:ncol(Rsim.output$out_Biomass)]
    n <- ncol(biomass)
    start.bio <- biomass[1, ]
    start.bio[which(start.bio == 0)] <- 1
    rel.bio <- matrix(NA, dim(biomass)[1], dim(biomass)[2])
    for(isp in 1:n) rel.bio[, isp] <- biomass[, isp] / start.bio[isp]
  }
  rel.bio
} 

names_spp <- run1$params$spname[2:94]

#Need to make z into data frame
z_1 <-rsim.plot.fun(run1)
z_1<-as.data.frame(z_1)
names(z_1) <- names_spp
z_1$month <- 1:600
data_1 <- gather(z_1, key = spp, value = biomass, -month )

#Plot functional group biomass over the 50-year simulation
#Graph with commercially valuable predators
plot1_comm <- data_1 %>% filter(spp %in% commercial_predators) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= factor(spp))) + xlab("Month") + ylab("Relative biomass") +
scale_color_discrete(type = "viridis", labels= c("Greenstriped" = expression(bold("Greenstriped rockfish")), "P. Ocean Perch"= expression(bold("P. ocean perch")), "Albacore" = expression(bold("Albacore")), "Arrowtooth" = expression(bold("Arrowtooth flounder")), "Canary rock."= expression(bold("Canary rock.")), "Lingcod"= expression(bold("Lingcod")), "Longspine thorny."= expression(bold("Longspine thornyhead")), "Other cephalopod"= expression(bold("Other cephal.")), "Sablefish"= expression(bold("Sablefish")), "Salmon"= expression(bold("Salmon")), "Sharks"= expression(bold("Sharks")), "Shelf rock."= expression(bold("Shelf rockfish")), "Shortspine thorny."= expression(bold("Shortspine thornyhead")), "Skates"= expression(bold("Skates")), "Slope rock."= expression(bold("Slope rockfish")), "Splitnose rockfish"= expression(bold("Splitnose rockfish")), "Widow rock."= expression(bold("Widow rockfish")), "Yellowtail rock."= expression(bold("Yellowtail rockfish")))) + 
  theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + labs(colour= "Commercial predators", title= "A") + theme(legend.text.align= 0, legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) + ggsave(width = 6, height = 6, "~/Documents/OTZ.manuscript/Images/plot.commercial.50.jpeg")
plot1_comm    

#Graph with noncommercially valuable predators
new_noncomm <- noncommerical_predators[- c(9,7,11,4,6,17,1,12,8,10,19,5,16,18,15,14,13,3)] #Remove predators that have commercial value as well 
plot1_noncomm <- data_1 %>% filter(spp %in% new_noncomm) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= factor(spp))) + xlab("Month") + ylab("Relative biomass") + 
scale_color_discrete(type = "viridis", labels= c("Western Gull" = "Western gull", "Tufted Puffin" = "Tufted puffin", "Sea Lions"= "Sea lions", "Resident Orcas"= "Resident orcas", "Leach's S. Petrel" ="Leach's s. petrel", "Juv. Ele. Seal"="Juv. elephant seal", "Double corm."= "Double cormorant", "Brandt's corm." = "Brandt's cormorant", "Juv. rock." = "Juv. rockfish")) + 
  theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + labs(colour= "Non-commercial predators", title= "B") + theme(legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) + ggsave(width = 6, height = 6, "~/Documents/OTZ.manuscript/Images/plot.noncommercial.50.jpeg")
plot1_noncomm    

#Graph for marine mammals
plot1_mammals <- data_1 %>% filter(spp %in% mammals) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= factor(spp))) + xlab("Month") + ylab("Relative biomass") + 
theme_bw() + 
scale_color_discrete(type = "viridis", labels= c("Adult N. Ele. Seal" = "Adult n. elephant seal", "Juv. Ele. Seal"= "Juv. elephant seal", "Resident Orcas"= "Resident orcas", "Sea Lions"= "Sea lions", "Transient Orcas"= "Transient orcas")) + 
theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + labs(colour = "Marine mammals", title= "A") + theme( legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12))
ggsave(width = 6, height = 5, "~/Documents/OTZ.manuscript/Images/plot.mammals.50.jpeg")
plot1_mammals

#Seperates legend and saves it as pdf document 
legend <- get_legend(plot1_mammals)
pdf("/Users/SallyDowd/Documents/OTZ.manuscript/Images/Legends/mammals.legend.pdf")
grid.draw(legend)
dev.off()

#Graph for species with highest economic value, NEED to change species names?
plot1_high.price <- data_1 %>% filter(spp %in% high.price) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= factor(spp))) + xlab("Month") + ylab("Relative biomass") + 
theme_bw() + 
scale_color_discrete(type= "viridis", labels= c("Albacore", "Lingcod", "Sablefish", "Salmon", "Shelf rockfish", "Shortspine thornyhead")) + 
theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + labs(colour = "Commercial predators", title= "B") + theme(legend.position= "none", legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) 
#ggsave(width = 6, height = 5, "~/Documents/OTZ.manuscript/Images/plot.highprice.50.jpeg")
plot1_high.price

#Seperates legend and saves it as pdf document 
legend <- get_legend(plot1_high.price)
pdf("/Users/SallyDowd/Documents/OTZ.manuscript/Images/Legends/price.legend.pdf", width= 6, height= 5)
grid.draw(legend)
dev.off()
```

#Depletion level of 25%
```{r}
target_sp <- "mesopelagics"
target_F <- 0.25
names(target_F) <- target_sp 

library(data.table)
scene2 <- copy(base_sim_scene)
run2 <- rsim.run(scene2, method = 'AB', years=2001)
for (yr in 2002:2050){
  # stock assessment based on current biomass * error
  # use this if including two species: assessment <- end_biomass(run3)[target_sp] * (runif(2) + 0.5)   
assessment <- end_biomass(run2)[target_sp] * (runif(length(target_sp)) + 0.5) #argument of length 0 means input is of length zero 
assessment
  # convert target F * assessment to a total catch 
  catches <- target_F * assessment
  # apply each catch to each species
  for (sp in target_sp){
    scene2 <- adjust.fishing(scene2, "ForcedCatch", sp, yr, value=catches[sp])
  }
  # run 1 year
  run2 <- rsim.step(scene2, run2, method = 'AB', year.end=yr)  #Says in Rsim document on NOAA github that older versions of Ecosim use AB and new versions use 'RK4', default for Rpath is RK4 but we aren't using it??
}  
rsim.plot(run2,all_species)  
```
#Modify data

```{r}
names_spp <- run2$params$spname[2:94]

#Need to make z into data frame
z_2 <-rsim.plot.fun(run2)
z_2<-as.data.frame(z_2)
names(z_2) <- names_spp
z_2$month <- 1:600
data_2 <- gather(z_2, key = spp, value = biomass, -month )
```

#Change over time graphs 
```{r}
change_2_overtime_df <- function(species) {
  x = seq(0,1, length.out = length(species))
  color = viridis(n = length(species), option= "B")
  par(mar = c(6,12,2,4))
  specieyear2 = run2$annual_Biomass[1, species]
  specieyearlast = run2$annual_Biomass[length(yearlist), species]
  value <- (specieyearlast/specieyear2)
  cbind(species, value)
}

change_2_total <- change_2_overtime_df(all_species)
change_2_total <- as.data.frame(change_2_total)

#3.4: Diet over 0.05
change_2_1_df<- change_2_overtime_df(diet_over_0.05)
change_2_1_df <- as.data.frame(change_2_1_df)
change_2_1 <- change_2_1_df %>% mutate(value = round(as.numeric(paste(value)), 2), "result" = cut(value, 
                      breaks = c(-Inf, 0.99, 1.001, Inf),
                      labels = c("Decreasing", "Neither", "Increasing"),
                      right = TRUE)) %>% arrange(value)

#Change species labels? change text size 
change_2_1_plot <- change_2_1 %>% ggplot(aes(col = result)) + geom_point(aes(x=value, y=reorder(species, desc(species)))) + 
scale_colour_grey() + 
scale_y_discrete(labels= c("Tufted Puffin"= "Tufted puffin", "P. Ocean Perch"= "P. ocean perch", "Leach's S. Petrel"= "Leach's s. petrel", "Longspine thorny."= "Longspine thornyhead", "Shelf rock."= "Shelf rockfish", "Slope rock."= "Slope rockfish", "Yellowtail rock."= "Yellowtail rockfish")) + 
  geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) +
  xlim(0.75, 1.25) + 
xlab("Relative biomass") + ylab("Species with high diet proportion ") + labs(col = "End biomass", title= "A") + theme_bw() + theme(panel.border = element_blank()) +
theme(axis.line = element_line(colour = "black")) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
theme(legend.position= "none", legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) + 
ggsave(width = 6, height = 5, "~/Documents/OTZ.manuscript/Images/dietover.25.jpeg")
change_2_1_plot 

legend <- get_legend(change_2_1_plot)
pdf("/Users/SallyDowd/Documents/OTZ.manuscript/Images/Legends/diet.legend.pdf", width= 6, height= 5)
grid.draw(legend)
dev.off()

#3.5: Diet under 0.05 

change_2_2_df<- change_2_overtime_df(diet_under_0.05_2)
change_2_2_df <- as.data.frame(change_2_2_df)
change_2_2 <- change_2_2_df %>% mutate(value = round(as.numeric(paste(value)), 2), "result" = cut(value, 
                      breaks = c(-Inf, 0.99, 1.001, Inf),
                      labels = c("Decreasing", "Neither", "Increasing"),
                      right = TRUE)) %>% arrange(value)


change_2_2_plot <- change_2_2 %>% ggplot(aes(col= result)) + geom_point(aes(x= value, y=reorder(species, desc(species)))) + 
scale_colour_grey() +  
 scale_y_discrete(labels= c("Western Gull"= "Western gull", "Resident Orcas"= "Resident orcas", "Juv. Ele. Seal"= "Juv. elephant seal", "Greenstriped"= "Greenstriped rockfish", "Double corm."= "Double cormorant",    "Brandt's corm."= "Brandt's cormorant", "Arrowtooth"= "Arrowtooth flounder", "Canary rock."= "Canary rockfish", "Juv. rock."= "Juvenile rockfish", "Other cephal."= "Other cephalopod", "Shortspine thorny."= "Shortspine thornyhead", "Widow rock."= "Widow rockfish")) + 
  xlim(0.75, 1.25) + geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) + 
xlab("Relative biomass") + ylab("Species with low diet proportion") + 
labs(col = "End biomass", title= "A") +
theme_bw() +
theme(panel.border = element_blank()) +
theme(axis.line = element_line(colour = "black")) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
#theme(legend.position= "none") #if don't want legend
  theme(legend.position= "none", legend.title= element_text(size = 12), legend.text = element_text(size= 10), axis.text=element_text(size=10), axis.title=element_text(size=12)) + 
ggsave(width = 7, height = 5, "~/Documents/OTZ.manuscript/Images/dietunder.25.jpeg")
change_2_2_plot
```



#Running simple stats on all depletion levels
```{r}
#Find increase and decrease of all predators for scenarios 1-4
biomass_increase_100 <- data_3 %>% filter(month==600, biomass >1) %>% mutate(difference= abs(1-biomass), percent = abs(1-biomass)*100)
biomass_increase_100

biomass_decrease_100 <- data_3 %>% filter(month==600, biomass <1) %>% mutate(difference= (1-biomass), percent= (1-biomass)*100)
biomass_decrease_100

biomass_increase_80 <- data_4 %>% filter(month==600, biomass >1) %>% mutate(difference= abs(1-biomass), percent = abs(1-biomass)*100)
biomass_increase_80

biomass_decrease_80 <- data_4 %>% filter(month==600, biomass <1) %>% mutate(difference= (1-biomass), percent = (1-biomass)*100)
biomass_decrease_80

biomass_increase_50 <- data_5 %>% filter(month==600, biomass >1) %>% mutate(difference= abs(1-biomass), percent = abs(1-biomass)*100)
biomass_increase_50

biomass_decrease_50 <- data_5 %>% filter(month==600, biomass <1) %>% mutate(difference= (1-biomass), percent= (1-biomass)*100)
biomass_decrease_50

biomass_increase_25 <- data_6 %>% filter(month==600, biomass >1) %>% mutate(difference = abs(1-biomass), percent = abs(1-biomass)*100)
biomass_increase_25

biomass_decrease_25 <- data_6 %>% filter(month==600, biomass < 1) %>% mutate(difference= (1-biomass), percent= (1-biomass)*100)
biomass_decrease_25

#Find species specific end biomass for 25% (and 50?) harvest rate 
predators_end2.1 <- data_2 %>% filter(spp %in% commercial_predators, month== 600)
predators_end2.1

predators_end2.2 <- data_2 %>% filter(spp %in% noncommerical_predators, month== 600)
predators_end2.2

predators_end2.3 <- data_2 %>% filter(spp %in% predators_mesopelagic, month == 600)
predators_end2.3

predators_end <- data_1 %>% filter(spp %in% commercial_predators, month== 600)
predators_end

predators_end2 <- data_1 %>% filter(spp %in% noncommerical_predators, month== 600)
predators_end2

predators_end3 <- data_1 %>% filter(spp %in% predators_mesopelagic, month == 600)
predators_end3

write.csv(predators_end, "/Users/SallyDowd/Documents/commercial.end.csv")
write.csv(predators_end2, "/Users/SallyDowd/Documents/noncommercial.end.csv")
write.csv(predators_end3, "/Users/SallyDowd/Documents/predators.end.csv")
```





