---
title: "Predator.increase.analysis"
output: html_document
---

#Code to explore the mainly positive and small biomass changes in most predators of mesopelagic fish shown in Rpath.main.results.Rmd for the paper titled "The economic tradeoffs and ecological implications associated with a potential mesopelagic fishery in the California Current". Code was created by Sally Dowd (sdowd20@berkeley.edu) and Melissa Chapman (mchapman@berkeley.edu).

#Load required pacakges
```{r}
library(dplyr)
library(readr)
library(Rpath)
library(tidyverse)
```

#Import original diet matrix (Diet_CalCur_USE.csv) and an unbalanced version of base parameters (Groupinfo_CalCur-USE.csv) from Koehn et al. (2016) and the Rsim results for the 25% and 50% harvest rate from Rpath.main.results.Rmd
```{r}
GroupInfo_CalCur_USE <- read_csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/GroupInfo_CalCur-USE.csv")
Diet <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Diet_CalCur_USE.csv")  
Diet <- Diet %>% rename("Juv. hake" = "Juv..hake", "Yellowtail rock."= "Yellowtail.rock.", "Shelf rock."= "Shelf.rock.", "P. Ocean Perch"= "P..Ocean.Perch", "Longspine thorny."= "Longspine.thorny.", "Tufted Puffin"= "Tufted.Puffin", "Leach's S. Petrel"= "Leach.s.S..Petrel","Fin whale"= "Fin.whale", "Slope rock."= "Slope.rock.", "carniv-zoops"= "carniv.zoops", "small jellies"= "small.jellies", "large jellies"= "large.jellies", "pandalid shp"= "pandalid.shp", "benthic shp"= "benthic.shp", "tanner crab"= "tanner.crb", "Benthic fish"= "Benthic.fish", "surf perch"= "surf.perch", "Other cephal."= "Other.cephal.", "Market squid"= "Market.squid", "Pacific Mackerel"= "Pacific.Mackerel")

scenario_7 <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Scenario_7.csv")
scenario_6 <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Scenario_6.csv")
```

#Webplot highlighting predator and prey connections to mesopelagic fish 
```{r}
jpeg(file="~/Documents/OTZ.Manuscript/Images/Sub.2/web.plot.jpeg",
width=300, height=525)
webplot(Cal_bal, eco.name= attr(Cal_bal, "California Current"), line.col="grey", labels=TRUE, highlight= "mesopelagics", highlight.col= c("blue", "red", "NA"), label.num=TRUE, label.cex=0.7, fleets=FALSE)

tiff("~/Documents/OTZ.Manuscript/Images/Sub.2/web.plot.jpeg", units="in", width=6, height=5, res=600, compression = 'lzw')
webplot(Cal_bal, eco.name= attr(Cal_bal, "California Current"), line.col="grey", labels=FALSE, highlight= "mesopelagics", highlight.col= c("blue", "red", "NA"), label.num=TRUE, label.cex=0.7, fleets=FALSE) + standard_theme
dev.off()
```

#Graph with weighted aggregate biomass for predators of mesopelagic fish, other prey of predators of mesopelagic fish and mesopelagic fish themselves. This code created Figure 4 in the paper. 
```{r}
#Define a common ggplot theme 
  standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour =
  "black", size=1)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.text.align= 0, legend.title= element_text(size = 10), legend.text =
  element_text(size= 9), axis.text=element_text(size=9),
  axis.title=element_text(size=10))
   
#Establish groupings of functional groups 
  predators_of_mesopelagic <- Diet[,c(1,16,19, 29:30, 34, 38:39, 43:49, 51:57, 59:61, 64,
  66:68, 70:71, 73:74, 77:81, 83, 86:89)]
  #All predators of mesopelagic fish

  predators_of_mesopelagic_0.05 <- predators_of_mesopelagic %>% pivot_longer(-DIETS) %>%
  filter(value >= 0.05) %>% rename(spp= DIETS) %>% mutate(classification= "Predators of
  mesopelagics") %>% select(-value, -name) #Predators of mesopelagic fish that have at     least 5% of their diet consist of mesopelagics

  prey_of_mesopelagic_preds <- Diet[c(3,6:10,18), c(1,30,34,39, 43, 49, 52, 55, 57, 64,
  71,77,86)] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% rename(spp= DIETS)%>%
  mutate(classification= "Prey of mesopelagics") %>% select(-value, -name) #Identify       prey of mesopelagics where higher diet proportion predators of mesopelagic fish have at   least 0.05 in their diet 

  pred_mesopelagic_prey_0.05 <- predators_of_mesopelagic %>% pivot_longer(-DIETS) %>%
  filter(value >= 0.05) %>% select(DIETS) %>% unique() %>% rename(spp = DIETS) %>%
  mutate(classification= "Other prey of predators of mesopelagics")
  #pred_mesopelagic_prey_0.05 is made up of the predators of mesopelagic fish (as defined above) that have other prey consisting of at least 5% of their diet    

  combo <- rbind(prey_of_mesopelagic_preds[!duplicated(prey_of_mesopelagic_preds$spp),],
  predators_mesopelagic_0.05,
  pred_mesopelagic_prey_0.05[!duplicated(pred_mesopelagic_prey_0.05$spp),])

  GroupInfo_edt <- GroupInfo_CalCur_USE %>% select(Group, Biomass) %>% rename(spp= Group)
  #select original biomass from group info to join with the predators (consumers) in both   groupings of functional groups

#Calculate weighted aggregated biomass for scenario 6 for predators of mesopelagic fish  and other prey of predators of mesopelagic fish.
  weighted_biomass <- scenario_6 %>% left_join(GroupInfo_edt, by= "spp") %>% left_join(combo,
  by= "spp") %>% drop_na() %>% group_by(classification, month) %>% mutate(total_biomass=
  sum(unique(Biomass)))%>% #sum initial biomass of each unique species groupings by month
  ungroup() %>% mutate(percent_biomass= Biomass/total_biomass) %>% #divide initial         biomass for a species by total biomass for a species group
  mutate(weighted_biomass= biomass*percent_biomass) %>% #get weighted biomass by           multiplying relative biomass by percent biomass 
  group_by(classification,month) %>% summarize(agg_rel_biomass=
  sum(weighted_biomass))
  #get aggregated relative biomass for each species classification (functional group       grouping) by summing weighted biomass for each species classification for each month  
#Make weighted aggregated biomass graph 
  mesopelagics_only <- scenario_6 %>% filter(spp== "mesopelagics") %>%
  rename(Mesopelagics=
  "spp")
  preds_other_prey <- ggplot() + geom_line(data=weighted_biomass, aes(x=month,
  y=agg_rel_biomass, col= as.factor(classification))) + geom_line(data=mesopelagics_only,
  aes(x = month, y = biomass, col= "Mesopelagics")) + labs(col= "Species classification")
  + scale_x_continuous(name= "Year", breaks= c(0, 120, 240, 360, 480, 600), labels=
  c("2001", "2010", "2020", "2030", "2040", "2050")) + geom_hline(yintercept=1, linetype=
  "solid", color= "black", size= 0.75) + xlab("Month") + ylab("Relative biomass") + 
  scale_color_discrete(type = "viridis") + standard_theme # + ggsave(width = 6, height =   4, "/Users/sallydowd/Documents/OTZ.manuscript/Images/Sub.2/preds.otherprey.25.jpeg",     dpi=600)
  preds_other_prey                                    
```

#Calculations for weighted biomass diet proportions of mesopelagic predators (in this case those with high diet proprotion of mesopelagics), mesopelagic fish, and certain mesopelagic fish prey items. The outputs from these calculations went into Appendix S3: Figure 3b. 
```{r}
avg_pmp <- prey_of_mesopelagic_preds %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for predators of mesopelagic fish on mesopelagic prey (that at least one of these predators prey on for at least 5% of their diet)

avg_pmf <- predators_of_mesopelagic[15,] %>% pivot_longer(-DIETS) %>% filter(value >= 0.05) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for mesopelagic predators on mesopelagic fish

avg_mfmp <- Diet[c(3,6,7, 8:10, 18), c(1,16)] %>% summarize(sum(mesopelagics))
#non-weighted value (just sum) of diet composition for mesopelagic fish on mesopelagic prey (if divided Biomass of mesopelagics by total_biomass of mesopelagics it would just equal 1)

avg_mpp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for select mesopelagic prey items on mesopelagic predators

avg_mfp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,16)] %>% summarize(mean(mesopelagics))
#weighted value of diet composition for mesopelagic fish on mesopelagic predators

avg_mpmf <- Diet[15, c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight))
#weighted value of diet composition for select mesopelagic prey items on mesopelagic fish

avgsmpmp <- Diet[c(3,6,7, 8:10, 18), c(1,4,7,8, 9:11, 19)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight)) 
#weighted value of diet composition for mesopelagic prey on mesopelagic prey  
  
avg_pp <- Diet[c(29,33,38, 42, 48, 51, 54, 56, 63, 70,76,85), c(1,30,34,39, 43, 49, 52, 55, 57, 64, 71,77,86)] %>% pivot_longer(-DIETS) %>% rename(spp= name) %>% left_join(GroupInfo_edt, by= "spp") %>% mutate(total_biomass= sum(unique(Biomass))) %>% mutate(bio_DC_weight = (Biomass*value)/total_biomass) %>% summarize(sum(bio_DC_weight)) 
#weighted value of diet composition for mesopelagic predators on mesopelagic predators
```
