---
title: "Rpath.main.results"
output: html_document
---

#Code to simulate the ecosystem-wide impacts of harvesting mesopelagic fishes at a yearly rate of 25 and 50% in the California Current for the paper titled "The economic tradeoffs and ecological implications associated with a potential mesopelagic fishery in the California Current". Code was originally created by Kerim Aydin (kerim.aydin@noaa.gov), modified by Laura Koehn (laura.koehn216@gmail.com), Sally Dowd (sdowd20@berkeley.edu), and Melissa Chapman (mchapman@berkeley.edu) 

#Load required packages
```{r}
library(Rpath)
library(data.table)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(tidyverse)
library(viridis)
```

#Import original diet matrix (Diet_CalCur_USE.csv) and an unbalanced version of base parameters (Groupinfo_CalCur-USE.csv) from Koehn et al. (2016). 
```{r}
#Define code source for base and diet parameters
Cdiet <- "../Data/Diet_CalCur_USE.csv"  #Diet matrix
Cbase <- "../Data/Groupinfo_CalCur-USE.csv"  #Base biomass, production, fishing, etc.

#Load unbalanced Ecopath model
CalCur_unbal <- read.rpath.params(Cbase, Cdiet)

#Read in created parameter files
check.rpath.params(CalCur_unbal)
```

#Run Ecopath in R through Rpath
#Solve for EE to balance an unbalanced Ecopath object and convert the Ecopath model to an Rpath object 
```{r}
#Balance the scenario using rpath function 
Cal_bal <- rpath(CalCur_unbal) 
Cal_bal

#Combine two data frames into a single data frame 
cbind(Cal_bal$Group, Cal_bal$EE)
cbind(Cal_bal$Group, Cal_bal$TL) 
```

#Define different sets of functional groups
```{r}
all_species <- Cal_bal$Group[1:92] 
mammals <- Cal_bal$Group[71:85] 
seabirds <- c(Cal_bal$Group[60:70], Cal_bal$Group[86:92])
pred_fish  <- Cal_bal$Group[33:59] 
Plankton <- c(Cal_bal$Group[1], Cal_bal$Group[5:10])
forage <- Cal_bal$Group[19:27]

#All predators of mesopelagic fish 
predators_mesopelagic <- c(Cal_bal$Group[15], Cal_bal$Group[18], Cal_bal$Group[28:29], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[42:48], Cal_bal$Group[50:56], Cal_bal$Group[58:60], Cal_bal$Group[63], Cal_bal$Group[65:67], Cal_bal$Group[69:70], Cal_bal$Group[72:73], Cal_bal$Group[76:80], Cal_bal$Group[82], Cal_bal$Group[85:88])
predators_mesopelagic

#Commercially valuable predators of mesopelagic fish
commercial_predators <- c(Cal_bal$Group[18], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[42:48], Cal_bal$Group[50:56], Cal_bal$Group[58:59])
commercial_predators

#Non-commercially valuable predators of mesopelagic fish
noncommerical_predators <- c(Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[33], Cal_bal$Group[37:38], Cal_bal$Group[43:47], Cal_bal$Group[50:56], Cal_bal$Group[58:60], Cal_bal$Group[63], Cal_bal$Group[65:67], Cal_bal$Group[69:70], Cal_bal$Group[72:73], Cal_bal$Group[76:80], Cal_bal$Group[82], Cal_bal$Group[85:88])
noncommerical_predators

#Predators that have a diet proportion of mesopelagic fish < 0.05
diet_under_0.05 <- c(Cal_bal$Group[15], Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[37], Cal_bal$Group[43:47], Cal_bal$Group[50], Cal_bal$Group[52:53], Cal_bal$Group[55], Cal_bal$Group[58:59], Cal_bal$Group[60], Cal_bal$Group[65:67], Cal_bal$Group[69],Cal_bal$Group[72:73], Cal_bal$Group[77:80], Cal_bal$Group[82], Cal_bal$Group[86:88]) #excludes mesopelagic fish group

diet_under_0.05_2 <- c(Cal_bal$Group[15], Cal_bal$Group[18], Cal_bal$Group[28], Cal_bal$Group[37], Cal_bal$Group[43:47], Cal_bal$Group[50], Cal_bal$Group[52:53], Cal_bal$Group[55], Cal_bal$Group[58:59], Cal_bal$Group[60], Cal_bal$Group[65:67], Cal_bal$Group[69],Cal_bal$Group[72:73], Cal_bal$Group[77:80], Cal_bal$Group[82], Cal_bal$Group[86:88]) #includes mesopelagic fish group 

#Predators that have a diet proportion of mesopelagic fish > 0.05
diet_over_0.05 <- c(Cal_bal$Group[29], Cal_bal$Group[33], Cal_bal$Group[38], Cal_bal$Group[42], Cal_bal$Group[48], Cal_bal$Group[51], Cal_bal$Group[54], Cal_bal$Group[56], Cal_bal$Group[63], Cal_bal$Group[70], Cal_bal$Group[76], Cal_bal$Group[85])
diet_over_0.05
```

#Initialize a base scenario
```{r}
yearlist = 2001:2050 #year range for simulations  

base_sim_scene <- rsim.scenario(Cal_bal, CalCur_unbal, years = yearlist)
scene0 <- base_sim_scene
run0 <- rsim.run(scene0, method='AB', years=yearlist) #base scenario created
rsim.plot(run0,all_species)
```

#Function to find the end biomass of a scenario in year 2050 (simulation month 600)
```{r}
end_biomass <- function(rsim){return(rsim$out_Biomass[dim(rsim$out_Biomass)[1], 2:(dim(rsim$out_Biomass)[2])])}
end_biomass
```

#Scenario 7: Apply a yearly 50% mesopelagic fish harvest rate to the base Rsim scenario and project the simulation forward for fifty years
```{r}
target_sp <- "mesopelagics" #Define target species
target_F <- 0.50 #Define target harvest rate 
names(target_F) <- target_sp 

#For year 2001 apply a 50% harvest rate to the original mesopelagic fish biomass. After year 1, and for every year through 2050, pause the operating model and assess the end biomass of that year for mesopelagic fish. Calculate the catch by multiplying the end biomass by an error of 0.5 and then force this catch for the next year. 
scene1 <- copy(base_sim_scene)
run1 <- rsim.run(scene1, method = 'AB', years=2001)
for (yr in 2002:2050){
  #Perform a stock assessment based on current biomass*error
  assessment <- end_biomass(run1)[target_sp] * (runif(length(target_sp)) + 0.5)
  assessment
  #Calculate total catch by multiplying target harvest rate by assessment 
  catches <- target_F * assessment
  #Apply the calculated catch to mesopelagic fish
  for (sp in target_sp){
    scene1 <- adjust.fishing(scene1, "ForcedCatch", sp, yr, value=catches[sp])
  }
    #Run 1 year
    run1 <- rsim.step(scene1, run1, method='AB', year.end=yr) 
    }  
rsim.plot(run1,all_species) 

#Change dataset from an Rpath object to a dataframe in order to plot with ggplot2 
rsim.plot.fun <- function(Rsim.output, spname, indplot = F, ...){
  opar <- par(no.readonly = T)
  if(indplot == F){
    biomass <- Rsim.output$out_Biomass[, 2:ncol(Rsim.output$out_Biomass)]
    n <- ncol(biomass)
    start.bio <- biomass[1, ]
    start.bio[which(start.bio == 0)] <- 1
    rel.bio <- matrix(NA, dim(biomass)[1], dim(biomass)[2])
    for(isp in 1:n) rel.bio[, isp] <- biomass[, isp] / start.bio[isp]
  }
  rel.bio
} 

names_spp <- run1$params$spname[2:94] 
z_1 <-rsim.plot.fun(run1)
z_1<-as.data.frame(z_1)
names(z_1) <- names_spp
z_1$month <- 1:600
data_1 <- gather(z_1, key = spp, value = biomass, -month )

write.csv(data_1, "/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Scenario.7.csv") 
```

#Scenario 6: Apply a yearly 25% mesopelagic fish harvest rate to the base Rsim scenario and project the simulation forward for fifty years
```{r}
target_sp <- "mesopelagics"
target_F <- 0.25
names(target_F) <- target_sp 

scene2 <- copy(base_sim_scene)
run2 <- rsim.run(scene2, method = 'AB', years=2001)
for (yr in 2002:2050){
  #Perform a stock assessment based on current biomass*error
  assessment <- end_biomass(run2)[target_sp] * (runif(length(target_sp)) + 0.5)
  assessment
  #Calculate total catch by multiplying target harvest rate by assessment 
  catches <- target_F * assessment
  #Apply the calculated catch to mesopelagic fish
  for (sp in target_sp){
    scene2 <- adjust.fishing(scene2, "ForcedCatch", sp, yr, value=catches[sp])
  }
    #Run 1 year
    run2 <- rsim.step(scene2, run2, method = 'AB', year.end=yr)  
    }  
rsim.plot(run2,all_species)  

#Change dataset from an Rpath object to a dataframe in order to plot with ggplot2 
names_spp <- run2$params$spname[2:94]
z_2 <-rsim.plot.fun(run2)
z_2<-as.data.frame(z_2)
names(z_2) <- names_spp
z_2$month <- 1:600
data_2 <- gather(z_2, key = spp, value = biomass, -month )
data_2 %>% filter(spp== "mesopelagics")

write.csv(data_2, "/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Scenario.6.csv")
```

#Use a function to create a change over time dataset for both the yearly 25% and 50% harvest rate on mesopelagic fish. This dataset has the relative end biomass (at the end of year 2050) of a functional group as the proportion of the original biomass of that group. 
```{r}
#50% harvest rate 
  change_1_overtime_df <- function(species) {
    x = seq(0,1, length.out = length(species))
    color = viridis(n = length(species), option= "B")
    par(mar = c(6,12,2,4))
    specieyear1 = run1$annual_Biomass[1, species]
    specieyearlast = run1$annual_Biomass[length(yearlist), species]
    value <- (specieyearlast/specieyear1)
    cbind(species, value)
  }
  change_1_total <- change_1_overtime_df(all_species)
  change_1_total <- as.data.frame(change_1_total)
  
#25% harvest rate
  change_2_overtime_df <- function(species) {
    x = seq(0,1, length.out = length(species))
    color = viridis(n = length(species), option= "B")
    par(mar = c(6,12,2,4))
    specieyear2 = run2$annual_Biomass[1, species]
    specieyearlast = run2$annual_Biomass[length(yearlist), species]
    value <- (specieyearlast/specieyear2)
    cbind(species, value)
  }
  change_2_total <- change_2_overtime_df(all_species)
  change_2_total <- as.data.frame(change_2_total)
```

#Change over time graphs for scenario 6 and 7 for all functional groups (with the exception of mesopelagic fish) to create Figure 3a-b in the paper and Appendix S3: Figure S2 in the supplementary material. The relative end biomass on the x axis the biomass at the end of the simulation in 2050 (month 600).  
```{r}
#Define a common ggplot theme
standard_theme <- theme_bw() + theme(panel.border = element_rect(fill=NA, colour = "black", size=1)) + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.text.align= 0, legend.title= element_text(size = 10), legend.text = element_text(size= 9), axis.text=element_text(size=9), axis.title=element_text(size=10))

#Scenario 7: yearly 50% harvest rate applied to mesopelagic fish 

  #Generate graph with functional groups with high diet dependence on mesopelagic fish:    Diet over 0.05 (mesopelagic fish make up at least 5% of their diet)
    change_1_1_df<- change_1_overtime_df(diet_over_0.05)
    change_1_1_df <- as.data.frame(change_1_1_df)
    change_1_1 <- change_1_1_df %>% mutate(value = round(as.numeric(paste(value)), 2),
    "result" = cut(value, breaks = c(-Inf, 0.99, 1.001, Inf), labels = c("Decreasing",
    "Neither", "Increasing"), right = TRUE)) %>% arrange(value)

    Predator_value <- read_csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Da
    ta/Predator.value.csv") #Load in economic value of functional groups
    change_1_1 <- change_1_1 %>% rename(biomass_prop= value) %>%    
    left_join(Predator_value, by="species")
    change_1_1[9,4] <- "Neither" #Make sure value column for juvenile hake says neither 
    change_1_1$Value <- factor(change_1_1$Value, levels = c("Commercial", 
    "Non-commercial", "Both", "Neither"))

    change_1_1_plot <- change_1_1 %>% ggplot(aes(col = Value)) +   
    geom_point(aes(x=biomass_prop, y=reorder(species, desc(species)))) + 
    scale_color_manual(values = c("#FF0000","#0000FF", "#00FF70", "#FF00FF")) + 
    scale_y_discrete(labels= c("Tufted Puffin"= "Tufted puffin", "P. Ocean Perch"= "P.
    ocean perch", "Leach's S. Petrel"= "Leach's s. petrel", "Longspine thorny."=
    "Longspine thornyhead", "Shelf rock."= "Shelf rockfish", "Slope rock."= "Slope
    rockfish", "Yellowtail rock."= "Yellowtail rockfish")) + 
    geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) +
    xlim(0.70, 1.35) + 
    xlab("Relative biomass") + ylab("Species with high diet proportion ") + labs(col =
    "Classification") + standard_theme
    change_1_1_plot 

  #Generate graph with functional groups with low diet dependence on mesopelagic fish:     Diet under 0.05 (mesopelagic fish make up less than 5% of their diet)
    change_1_2_df <- change_1_overtime_df(diet_under_0.05)
    change_1_2_df <- as.data.frame(change_1_2_df)
    change_1_2 <- change_1_2_df %>% mutate(value = round(as.numeric(paste(value)), 2),
    "result" = cut(value, breaks = c(-Inf, 0.99, 1.001, Inf), labels = c("Decreasing",
    "Neither", "Increasing"), right = TRUE)) %>% arrange(value)
    change_1_2 <- change_1_2 %>% rename(biomass_prop= value) %>%
    left_join(Predator_value, by="species") 
    change_1_2[1,4] <- "Neither"
    change_1_2 <- change_1_2[-1,] #Remove mesopelagic fish 
    change_1_2$Value <- factor(change_1_2$Value, levels = c("Non-commercial", "Both",
    "Neither"))

    change_1_2_plot <- change_1_2 %>% ggplot(aes(col= Value)) + geom_point(aes(x=
    biomass_prop, y=reorder(species, desc(species)))) + 
    scale_color_manual(values = c("#0000FF", "#00FF70")) +
    scale_y_discrete(labels= c("Western Gull"= "Western gull", "Resident Orcas"=
    "Resident orcas", "Juv. Ele. Seal"= "Juv. n. elephant seal", "Greenstriped"=
    "Greenstriped rockfish", "Double corm."= "Double cormorant", "Brandt's corm."=
    "Brandt's cormorant", "Arrowtooth"= "Arrowtooth flounder", "Canary rock."= "Canary
    rockfish", "Juv. rock."= "Juv. rockfish", "Other cephal."= "Other cephalopod",
    "Shortspine thorny."= "Shortspine thornyhead", "Widow rock."= "Widow rockfish",
    "mesopelagics"= "Mesopelagics", "Sea Lions"= "Sea lions", "Splitnose rock."=
    "Splitnose rockfish")) + 
    xlim(0.70, 1.35) + geom_vline(xintercept=1, linetype= "solid", color= "black", size=
    0.75) + xlab("Relative biomass") + ylab("Species with low diet proportion") + 
    labs(col = "Classification") + standard_theme
    change_1_2_plot

  #Generate graph with functional groups with no diet dependence on mesopelagic fish.      This means that these functional groups have 0% of their diet consist of mesopelagic     fish. 
    commercial_predators_df <- as.data.frame(commercial_predators) %>% rename(species=
    commercial_predators)
    noncommercial_predators_df <- as.data.frame(noncommerical_predators) %>%
    rename(species= noncommerical_predators)
    combo <- rbind(commercial_predators_df, noncommercial_predators_df, by= "species")       %>% mutate(Dependence= 1)
    no_dependence <- change_1_total %>% left_join(combo, by= "species")
    no_dependence[is.na(no_dependence)] <- 0 
    no_dependence <-no_dependence[!duplicated(no_dependence$species),] 
    no_dependence <- no_dependence[-c(15,30),] #Need to remove mesopelagic fish and          juvenile hake from dataset as they don't have commercial or non-commercial value so      are included here but they do have diet dependence on mesopelagic fish
    no_dependence_final <- no_dependence %>% filter(Dependence==0)
    no_dependence_final <- no_dependence_final %>% mutate(value =
    round(as.numeric(paste(value)), 2), "result" = cut(value, breaks = c(-Inf, 0.99,
    1.001, Inf), labels = c("Decreasing", "Neither", "Increasing"), right = TRUE)) %>%
    arrange(value)

    change_1_3_df <- no_dependence_final
    change_1_3 <- change_1_3_df %>% mutate(value = round(as.numeric(paste(value)), 2),
    "result" = cut(value, breaks = c(-Inf, 0.99, 1.001, Inf), labels = c("Decreasing",
    "Neither", "Increasing"), right = TRUE)) %>% arrange(value) %>% rename(biomass_prop=
    value) 
    change_1_3_plot <- change_1_3 %>% ggplot() + geom_point(aes(x= biomass_prop,
    y=reorder(species, desc(species)))) + 
    xlim(0.70, 1.35) + geom_vline(xintercept=1, linetype= "solid", color= "black", size=
    0.75) + 
    xlab("Relative biomass") + ylab("Species with no diet proportion") + 
    scale_y_discrete(labels= c("Adult N. Ele. Seal"= "Adult n. elephant seal",
    "amphipods"= "Amphipods", "benthic shp"= "Benthic shrimp", "Brown Pelican"= "Brown
    pelican", "Cali. gull"= "California gull", "carniv-zoops" = "Carnivorous
    zooplankton", "copepods"= "Copepods", "Dungeness"= "Dungeness crab", "epibenthic"=
    "Epibenthic", "euphausiids"= "Euphausiids", "infauna"= "Infauna", "Juv. round."=
    "Juv. roundfish", "Juv. thorny."= "Juv. thornyhead", "large jellies"= "Large
    jellies", "micro-zoop"= "Microzooplankton", "Nearshore rock."= "Nearshore rockfish",
    "Pacific Mackerel"= "Pacific mackerel", "pandalid shp"= "Pandalid shrimp",
    "phytoplankton"= "Phytoplankton", "Pigeon Guillemot"= "Pigeon guillemot", "Rhino.
    auk"= "Rhinoceros auklet", "small jellies"= "Small jellies", "surf perch"= "Surf
    perch", "tanner crb"= "Tanner crab", "Transient Orcas"= "Transient orcas", "Yelloweye
    rock."= "Yelloweye rockfish")) + labs(col = "Classification") + standard_theme
    change_1_3_plot

#Scenario 6: yearly 25% harveset rate applied to mesopelagics
    
 #Generate graph with functional groups with high diet dependence on mesopelagic fish:    Diet over 0.05 (mesopelagic fish make up at least 5% of their diet)
    change_2_1_df <- change_2_overtime_df(diet_over_0.05)
    change_2_1_df <- as.data.frame(change_2_1_df)
    change_2_1 <- change_2_1_df %>% mutate(value = round(as.numeric(paste(value)), 2),
    "result" = cut(value, breaks = c(-Inf, 0.99, 1.001, Inf), labels = c("Decreasing",
    "Neither", "Increasing"), right = TRUE)) %>% arrange(value)
    
    change_2_1 <- change_2_1 %>% rename(biomass_prop= value) %>%        
    left_join(Predator_value, by="species")
    change_2_1[10,4] <- "Neither"
    change_2_1$Value <- factor(change_2_1$Value, levels = c("Commercial",
    "Non-commercial", "Both", "Neither"))
    change_2_1_plot <- change_2_1 %>% ggplot(aes(col = Value)) +     
    geom_point(aes(x=biomass_prop, y=reorder(species, desc(species)))) + 
    scale_color_manual(values = c("#FF0000","#0000FF", "#00FF70", "#FF00FF")) + 
    scale_y_discrete(labels= c("Tufted Puffin"= "Tufted puffin", "P. Ocean Perch"= "P.
    ocean perch", "Leach's S. Petrel"= "Leach's s. petrel", "Longspine thorny."=
    "Longspine thornyhead", "Shelf rock."= "Shelf rockfish", "Slope rock."= "Slope
    rockfish", "Yellowtail rock."= "Yellowtail rockfish")) + 
    geom_vline(xintercept=1, linetype= "solid", color= "black", size= 0.75) +
    xlim(0.70, 1.35) + 
    xlab("Relative biomass") + ylab("Species with high diet proportion ") + labs(col =
    "Classification") + standard_theme 
    change_2_1_plot 

  #Generate graph with functional groups with low diet dependence on mesopelagic fish:     Diet under 0.05 (mesopelagic fish make up less than 5% of their diet)
    change_2_2_df <- change_2_overtime_df(diet_under_0.05_2)
    change_2_2_df <- as.data.frame(change_2_2_df)
    change_2_2 <- change_2_2_df %>% mutate(value = round(as.numeric(paste(value)), 2),
    "result" = cut(value, breaks = c(-Inf, 0.99, 1.001, Inf), labels = c("Decreasing",
    "Neither", "Increasing"), right = TRUE)) %>% arrange(value)
    change_2_2 <- change_2_2 %>% rename(biomass_prop= value) %>%
    left_join(Predator_value, by="species") 
    change_2_2[1,4] <- "Neither"
    change_2_2 <- change_2_2[-1,]
    change_2_2$Value <- factor(change_2_2$Value, levels = c("Non-commercial", "Both",
    "Neither"))

    change_2_2_plot <- change_2_2 %>% ggplot(aes(col= Value)) + geom_point(aes(x=
    biomass_prop, y=reorder(species, desc(species)))) + 
    scale_color_manual(values = c("#0000FF", "#00FF70")) +
    scale_y_discrete(labels= c("Western Gull"= "Western gull", "Resident Orcas"=
    "Resident orcas", "Juv. Ele. Seal"= "Juv. n. elephant seal", "Greenstriped"=
    "Greenstriped rockfish", "Double corm."= "Double cormorant", "Brandt's corm."=
    "Brandt's cormorant", "Arrowtooth"= "Arrowtooth flounder", "Canary rock."= "Canary
    rockfish", "Juv. rock."= "Juv. rockfish", "Other cephal."= "Other cephalopod",
    "Shortspine thorny."= "Shortspine thornyhead", "Widow rock."= "Widow rockfish",
    "mesopelagics"= "Mesopelagics", "Sea Lions"= "Sea lions", "Splitnose rock."=
    "Splitnose rockfish")) + 
    xlim(0.75, 1.35) + geom_vline(xintercept=1, linetype= "solid", color= "black",
    size= 0.75) + 
    xlab("Relative biomass") + ylab("Species with low diet proportion") + 
    labs(col = "Classification") + standard_theme 
    change_2_2_plot

  #Generate graph with functional groups with no diet dependence on mesopelagic fish.      This means that these functional groups have 0% of their diet consist of mesopelagic     fish.
    no_dependence_2 <- change_2_total %>% left_join(combo, by= "species")
    no_dependence_2[is.na(no_dependence_2)] <- 0 #Functional groups with no dependence on     mesopelagics have a 0
    no_dependence_2 <-no_dependence_2[!duplicated(no_dependence_2$species),] 
    no_dependence_2 <- no_dependence_2[-c(15,30),] #Remove mesopelagics and juvenile         hake- dependence on mesopelagics but don't have commercial or non-commercial value so     will be weird with filtering
    no_dependence_final_2 <- no_dependence_2 %>% filter(Dependence==0)
    change_2_3_df <- no_dependence_final_2
    change_2_3 <- change_2_3_df %>% mutate(value = round(as.numeric(paste(value)), 2),
    "result" = cut(value, breaks = c(-Inf, 0.99, 1.001, Inf), labels = c("Decreasing",
    "Neither", "Increasing"), right = TRUE)) %>% arrange(value) %>% rename(biomass_prop=
    value) 

    change_2_3 %>% filter(result== "Neither")
    change_2_3_plot <- change_2_3 %>% ggplot() + geom_point(aes(x= biomass_prop,
    y=reorder(species, desc(species)))) + 
    xlim(0.70, 1.35) + geom_vline(xintercept=1, linetype= "solid", color= "black", size=
    0.75) + 
    xlab("Relative biomass") + ylab("Species with no diet proportion") + 
    scale_y_discrete(labels= c("Adult N. Ele. Seal"= "Adult n. elephant seal",
    "amphipods"= "Amphipods", "benthic shp"= "Benthic shrimp", "Brown Pelican"= "Brown
    pelican", "Cali. gull"= "California gull", "carniv-zoops" = "Carnivorous
    zooplankton", "copepods"= "Copepods", "Dungeness"= "Dungeness crab", "epibenthic"=
    "Epibenthic", "euphausiids"= "Euphausiids", "infauna"= "Infauna", "Juv. round."=
    "Juv. roundfish", "Juv. thorny."= "Juv. thornyhead", "large jellies"= "Large
    jellies", "micro-zoop"= "Microzooplankton", "Nearshore rock."= "Nearshore rockfish",
    "Pacific Mackerel"= "Pacific mackerel", "pandalid shp"= "Pandalid shrimp",
    "phytoplankton"= "Phytoplankton", "Pigeon Guillemot"= "Pigeon guillemot", "Rhino.
    auk"= "Rhinoceros auklet", "small jellies"= "Small jellies", "surf perch"= "Surf
    perch", "tanner crb"= "Tanner crab", "Transient Orcas"= "Transient orcas", "Yelloweye
    rock."= "Yelloweye rockfish")) + labs(col = "Classification") + standard_theme 
    change_2_3_plot
```

#Time series graphs for scenario 6 and 7 for functional groups with high and low diet dependence on mesoopelagic fish. This code created Appendix S3: Figure 1a-d in the supplementary material. The relative biomass on the y axis is the proportion of original biomass throughout the entire 50-year simulation. 
```{r}
scenario_7 <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Scenario.7.csv") #Results from 50% harvest rate 
scenario_6 <- read.csv("/Users/sallydowd/Documents/GitHub/mesopelagic-fishing/Data/Scenario.6..csv") #Results from 25% harvest rate

#Graph for functional groups with high diet dependence on mesopelagic fish
  Predator_value_ts <- Predator_value %>% rename(spp= "species")
  scenario_7_high <- scenario_7 %>% select(-X) %>% filter(spp %in% diet_over_0.05) %>% 
  left_join(Predator_value_ts, by= "spp")  
  scenario_7_high$Value <- factor(scenario_7_high$Value, levels = c("Commercial",
  "Non-commercial", "Both", "Neither"))
  scenario_7_high[1:600, 4] <- "Neither"

  scenario_6_high <- scenario_6 %>% select(-X) %>% filter(spp %in% diet_over_0.05) %>% 
  left_join(Predator_value_ts, by= "spp") 
  scenario_6_high$Value <- factor(scenario_6_high$Value, levels = c("Commercial",
  "Non-commercial", "Both", "Neither"))
  scenario_6_high[1:600, 4] <- "Neither"

  plot_scenario_7_high <- scenario_7_high %>% mutate(spp = fct_reorder(spp, biomass,
  tail, n = 1,
  .desc = TRUE)) %>% ggplot(aes(x= month, y = biomass)) + geom_line(aes(colour= spp,
  linetype= Value)) + 
  xlab("Month") + ylab("Relative biomass") + scale_x_continuous(name= "Year", breaks=
  c(0, 120, 240, 360, 480, 600), labels= c("2001", "2010", "2020", "2030", "2040",
  "2050")) + 
  scale_color_discrete(type = "viridis", labels= c("Greenstriped" = "Greenstriped
  rockfish", "P. Ocean Perch"= "P. ocean perch", "Albacore" = "Albacore", "Arrowtooth" =
  "Arrowtooth flounder", "Canary rock."= "Canary rockfish", "Lingcod"= "Lingcod",
  "Longspine thorny."= "Longspine thornyhead", "Shelf rock."= "Shelf rockfish",
  "Shortspine thorny."= "Shortspine thornyhead", "Slope rock."= "Slope rockfish", "Widow
  rock."= "Widow rockfish", "Yellowtail rock."= "Yellowtail rockfish")) + labs(colour=
  "Species with high diet proportion", title= "c") + standard_theme +           
  geom_hline(yintercept=1, linetype= "solid", color= "black", size= 0.75) +
  guides(linetype = guide_legend(order = 2),col = guide_legend(order = 1)) +
  scale_linetype_manual("Classification",values=c("Commercial"= 1,
  "Non-commercial"=2,"Both"=3, "Neither"=4)) + 
  ggsave(width = 8.75, height = 5.25,
  "~/Documents/OTZ.manuscript/Images/Sub.2/high.diet.50.jpeg", dpi=600)
  plot_scenario_7_high

  plot_scenario_6_high <- scenario_6_high %>% filter(spp %in% diet_over_0.05) %>%
  mutate(spp =
  fct_reorder(spp, biomass, tail, n = 1, .desc = TRUE)) %>% ggplot(aes(x= month, y =
  biomass)) + geom_line(aes(colour= spp, linetype= Value)) + 
  xlab("Month") + ylab("Relative biomass") + scale_x_continuous(name= "Year", breaks=
  c(0, 120, 240, 360, 480, 600), labels= c("2001", "2010", "2020", "2030", "2040",
  "2050")) + 
  scale_color_discrete(type = "viridis", labels= c("Greenstriped" = "Greenstriped
  rockfish", "P. Ocean Perch"= "P. ocean perch", "Albacore" = "Albacore", "Arrowtooth" =
  "Arrowtooth flounder", "Canary rock."= "Canary rockfish", "Lingcod"= "Lingcod",
  "Longspine thorny."= "Longspine thornyhead", "Shelf rock."= "Shelf rockfish",
  "Shortspine thorny."= "Shortspine thornyhead", "Slope rock."= "Slope rockfish", "Widow
  rock."= "Widow rockfish", "Yellowtail rock."= "Yellowtail rockfish", "Juv. Ele. Seal" =
  "Juv n. elephant seal")) + labs(colour= "Species with high diet proportion", title=
  "a") + labs(colour= "Species with high diet proportion") + standard_theme + 
  geom_hline(yintercept=1, linetype= "solid", color= "black", size= 0.75) +
  guides(linetype = guide_legend(order = 2),col = guide_legend(order = 1)) +
  scale_linetype_manual("Classification",values=c("Commercial"= 1,
  "Non-commercial"=2,"Both"=3, "Neither"=4)) + 
  ggsave(width = 8.75, height = 5.25, "~/Documents/OTZ.manuscript/Images/Sub.2/high.diet.25.jpeg", dpi=600)
  plot_scenario_6_high

#Graph for functional groups with low diet dependence on mesopelagic fish
  scenario_7_low <- scenario_7 %>% select(-X) %>% filter(spp %in% diet_under_0.05) %>%  
  left_join(Predator_value_ts, by= "spp") %>% filter(spp !="mesopelagics")
  scenario_7_low$Value <- factor(scenario_7_low$Value, levels = c("Commercial",
  "Non-commercial","Both", "Neither"))
  scenario_6_low <- scenario_6 %>% select(-X) %>% filter(spp %in% diet_under_0.05) %>%     left_join(Predator_value_ts, by= "spp") %>% filter(spp !="mesopelagics")
  scenario_6_low$Value <- factor(scenario_6_low$Value, levels = c("Commercial",
  "Non-commercial","Both", "Neither"))
  
   plot_scenario_7_low <- scenario_7_low %>% filter(spp %in% diet_under_0.05) %>%
   mutate(spp =
  fct_reorder(spp, biomass, tail, n = 1, .desc = TRUE)) %>% ggplot(aes(x= month, y =
  biomass)) + geom_line(aes(colour= spp, linetype= Value)) + 
  xlab("Month") + ylab("Relative biomass") + scale_x_continuous(name= "Year", breaks=
  c(0, 120, 240, 360, 480, 600), labels= c("2001", "2010", "2020", "2030", "2040",
  "2050")) + 
  scale_color_discrete(type = "viridis", labels= c("Greenstriped" = "Greenstriped
  rockfish", "P. Ocean Perch"= "P. ocean perch", "Arrowtooth" = "Arrowtooth flounder",
  "Canary rock."= "Canary rockfish", "Longspine thorny."= "Longspine thornyhead", "Shelf
  rock."= "Shelf rockfish", "Shortspine thorny."= "Shortspine thornyhead", "Slope rock."=
  "Slope rockfish", "Widow rock."= "Widow rockfish", "Yellowtail rock."= "Yellowtail
  rockfish", "Western Gull"= "Western gull", "Resident Orcas"= "Resident orcas", "Double
  corm."= "Double cormorant", "Brandt's corm."= "Brandt's cormorant", "Juv. rock."= "Juv.
  rockfish", "Other cephal."= "Other cephalopod", "Sea Lions"= "Sea lions", "Splitnose
  rock."= "Splitnose rockfish", "Juv. Ele. Seal"= "Juv. n. elephant seal")) +
  labs(colour= "Species with low diet proportion", title= "d") + standard_theme + 
  geom_hline(yintercept=1, linetype= "solid", color= "black", size= 0.75) +
  guides(linetype = guide_legend(order = 2),col = guide_legend(order = 1)) +
  scale_linetype_manual("Classification",values=c("Non-commercial"=2,"Both"=3)) + 
  ggsave(width = 8.75, height = 5.25, "~/Documents/OTZ.manuscript/Images/Sub.2/low.diet.50.jpeg", dpi=600)
  plot_scenario_7_low

  plot_scenario_6_low <- scenario_6_low %>% filter(spp %in% diet_under_0.05) %>%
  mutate(spp =
  fct_reorder(spp, biomass, tail, n = 1, .desc = TRUE)) %>% ggplot(aes(x= month, y =
  biomass)) + geom_line(aes(colour= spp, linetype= Value)) + 
  xlab("Month") + ylab("Relative biomass") + scale_x_continuous(name= "Year", breaks=
  c(0, 120, 240, 360, 480, 600), labels= c("2001", "2010", "2020", "2030", "2040",
  "2050")) + 
  scale_color_discrete(type = "viridis", labels= c("Greenstriped" = "Greenstriped
  rockfish", "P. Ocean Perch"= "P. ocean perch", "Arrowtooth" = "Arrowtooth flounder",  
  "Canary rock."= "Canary rockfish", "Longspine thorny."= "Longspine thornyhead", "Shelf
  rock."= "Shelf rockfish", "Shortspine thorny."= "Shortspine thornyhead", "Slope rock."=
  "Slope rockfish", "Widow rock."= "Widow rockfish", "Yellowtail rock."= "Yellowtail
  rockfish", "Western Gull"= "Western gull", "Resident Orcas"= "Resident orcas", "Double
  corm."= "Double cormorant", "Brandt's corm."= "Brandt's cormorant", "Juv. rock."= "Juv.
  rockfish", "Other cephal."= "Other cephalopod", "Sea Lions"= "Sea lions", "Splitnose
  rock."= "Splitnose rockfish", "Juv. Ele. Seal"= "Juv. n. elephant seal")) +
  labs(colour= "Species with low diet proportion", title= "b") + standard_theme +
  geom_hline(yintercept=1, linetype= "solid", color= "black", size= 0.75) +
  guides(linetype = guide_legend(order = 2),col = guide_legend(order = 1)) +
  scale_linetype_manual("Classification",values=c("Non-commercial"=2,"Both"=3)) +
  ggsave(width = 8.75, height = 5.25,
  "~/Documents/OTZ.manuscript/Images/Sub.2/low.diet.25.jpeg", dpi=600)
  plot_scenario_6_low
```

#Calculations for paper statistics.
```{r}
#Find which functional groups increased/decreased at the end of the 25% and 50% harvest rate scenarios 
  biomass_increase_25 <- scenario_6 %>% filter(month==600, biomass >1) %>%
  mutate(difference= abs(1-biomass), percent = abs(1-biomass)*100)
  biomass_increase_25
  
  biomass_decrease_25 <- scenario_6 %>% filter(month==600, biomass <1) %>%
  mutate(difference= (1-biomass), percent= (1-biomass)*100)
  biomass_decrease_25
  
  biomass_increase_50 <- scenario_7 %>% filter(month==600, biomass >1) %>%
  mutate(difference= abs(1-biomass), percent = abs(1-biomass)*100)
  biomass_increase_50
  
  biomass_decrease_50 <- scenario_7 %>% filter(month==600, biomass <1) %>%
  mutate(difference= (1-biomass), percent = (1-biomass)*100)
  biomass_decrease_50

#Find the relative end biomass for commercial and non-commercial predators of mesopelagic fish at the end of the 25% and 50% harvest rate 
  predators_end25_comm <- scenario_6 %>% filter(spp %in% commercial_predators, month==
  600) 
  predators_end25_nonc <- scenario_6 %>% filter(spp %in% noncommerical_predators, month==
  600)
  predators_end50_comm <- scenario_7 %>% filter(spp %in% commercial_predators, month==
  600)
  predators_end50_nonc <- scenario_7 %>% filter(spp %in% noncommerical_predators, month==
  600)
```
